---
title: "CAMS: CAMS_STATAREA_STOCK"
#author: "Ben Galuardi"
output:
  html_document: default
  github_document:
  pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(odbc)
library(ROracle)
library(dplyr, warn.conflicts = FALSE)

# # unlock keyring
# keyring::keyring_unlock("apsd_ma")
# 
# # local run
# # dw_apsd <- config::get(value = "apsd", file = "K:/R_DEV/config.yml")
# 
# # if on server..
# # dw_apsd <- config::get(value = "maps", file = "~/config.yml")
# 
# 
# # connect to MAPS
# 
# con_maps = apsdFuns::roracle_login(key_name = 'apsd_ma', key_service = 'maps')


```

```{r, include = FALSE, echo = FALSE}
Sys.setenv(TZ = "America/New_York")
Sys.setenv(ORA_SDTZ = "America/New_York")
  
if(file.exists(here::here("vignettes/articles", "pw.csv"))) {
  pw_df <- readr::read_csv(here::here("vignettes/articles", "pw.csv"), col_types = "c")
  pw <- pw_df$pw
} else {
  pw <- rstudioapi::askForPassword("Enter keyring password")
}

keyring::keyring_unlock("apsd_ma", password = pw) # will prompt for password
con_maps <- apsdFuns::roracle_login(key_name = 'apsd_ma', key_service = 'maps')
```

## Summary
This discard support table translates species stat areas to species stock areas used for discard stratification based on area stratifications used in current stock assessments and quota monitoring.  Species specific area stratifications have agreed upon by a discard working group.


## Description

- **OUTPUT TABLE:** CAMS_STATAREA_STOCK
- **YEARS:** N/A
- **RESOLUTION:** N/A
- **DEVELOPMENT LANGUAGE:** N/A
- **CODE:** https://github.com/NOAA-Fisheries-Greater-Atlantic-Region/discaRd/tree/model_estimator/CAMS

```{r table_flow0, echo=F,fig.height=3,fig.width=8, fig.cap="Figure 1.   CAMS_STATAREA_STOCK table lineage"}
DiagrammeR::mermaid("
  graph LR
  CAMS_STATAREA_STOCK --> GF_and_Non_GF_Discard_Modules
        ")
```

## Data Dictionary

```{r echo = F}
comments <- ROracle::dbGetQuery(con_maps, "SELECT * FROM all_col_comments WHERE owner = 'MAPS' AND table_name = 'CAMS_STATAREA_STOCK'")

comments %>%
dplyr::select(
COLUMN = COLUMN_NAME,
DESCRIPTION = COMMENTS) %>%
knitr::kable(booktabs = TRUE) %>%
kableExtra::kable_styling(bootstrap_options = c("striped", 
"hover", "condensed", "responsive"),full_width = FALSE,position = "left", fixed_thead = TRUE,
font_size = 10)
```


