---
title: "CAMS: GF_EM_DELTA_VTR_DISCARD"
#author: "Ben Galuardi"
output:
  html_document: default
  github_document:
  pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(odbc)
library(ROracle)
library(dplyr, warn.conflicts = FALSE)

# unlock keyring
keyring::keyring_unlock("apsd_ma")

# local run
# dw_apsd <- config::get(value = "apsd", file = "K:/R_DEV/config.yml")

# if on server..
# dw_apsd <- config::get(value = "maps", file = "~/config.yml")


# connect to MAPS

con_maps = apsdFuns::roracle_login(key_name = 'apsd_ma', key_service = 'maps')


```

## Summary
This table contains VTR discard estimates, EM discard estimates for EM reviewed GF trips and delta model estimates for unreviewed EM trips from FY20 forward. From May 1, 2021 to December 31, 2021 VTR estimates were used until adequate data was available to fit the delta model. Discards from observer coverage take precedent over discard estimates from EM. 


## Description

- **OUTPUT TABLE:** GF_EM_DELTA_VTR_DISCARD
- **YEARS:** 2020+
- **RESOLUTION:** Species discard estimate by VTRSERNO (subtrip)
- **DEVELOPMENT LANGUAGE:** R
- **CODE:** https://github.com/NOAA-Fisheries-Greater-Atlantic-Region/discaRd/tree/model_estimator/CAMS


## Data Sources
nemis.NEMIS_TRIP_MV@nova
nemis.NEMIS_HAUL_MV@nova
nemis.NEMIS_CATCH_MV@nova
noaa.document 
noaa.images 

```{r table_flow0, echo=F,fig.height=3,fig.width=8, fig.cap="Figure 1.   GF_EM_DELTA_VTR_DISCARD table lineage"}
DiagrammeR::mermaid("
  graph LR
nemis.NEMIS_TRIP_MV --> GF_EM_DELTA_VTR_DISCARD
nemis.NEMIS_HAUL_MV --> GF_EM_DELTA_VTR_DISCARD
nemis.NEMIS_CATCH_MV --> GF_EM_DELTA_VTR_DISCARD
noaa.document --> GF_EM_DELTA_VTR_DISCARD
noaa.images --> GF_EM_DELTA_VTR_DISCARD
        ")
```

## Approach
Criteria for EM data to be used to fit delta model:
(ct_tot >= 30 AND fishnk_pct >= .1) THEN 'N' ELSE haul_observed

Delta model:
 EM_DISCARD ~ LOG_EVTR_DISCARD + (1+LOG_EVTR_DISCARD|SPECIES_ID) + (1+LOG_EVTR_DISCARD | PERMIT)


## Data Dictionary

```{r echo = F}
comments <- ROracle::dbGetQuery(bcon, "SELECT * FROM all_col_comments WHERE owner = 'MAPS' AND table_name = 'GF_EM_DELTA_VTR_DISCARD'")

comments %>%
dplyr::select(
COLUMN = COLUMN_NAME,
DESCRIPTION = COMMENTS) %>%
knitr::kable(booktabs = TRUE) %>%
kableExtra::kable_styling(bootstrap_options = c("striped", 
"hover", "condensed", "responsive"),full_width = FALSE,position = "left", fixed_thead = TRUE,
font_size = 10)
```


