---
title: "CAMS: CAMS_STATAREA_STOCK"
#author: "Ben Galuardi"
output:
  html_document: default
  github_document:
  pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(odbc)
library(ROracle)
library(dplyr, warn.conflicts = FALSE)


 dw_maps <- config::get(config = "maps", file = "~/config_group.yml")
# 
 # bcon <- dbConnect(odbc::odbc(), 
 # 									DSN = dw_maps$dsn, 
 # 									UID = dw_maps$uid, 
 # 									PWD = dw_maps$pwd)

# Connect to database - move this to config file in the future - quick addition for server
  connectString <- paste(
    "(DESCRIPTION=",
    "(ADDRESS=(PROTOCOL=tcp)(HOST=", dw_maps$host, ")(PORT=", dw_maps$port, "))",
    "(CONNECT_DATA=(SERVICE_NAME=",dw_maps$svc, ")))",
    sep = ""
  )

# Connect to oracle each loop in case of timeouts
    bcon <- ROracle::dbConnect(
      drv = ROracle::Oracle(),
      username = dw_maps$uid,
      password = dw_maps$pwd,
      dbname = connectString
    )

```


## Summary
This discard support table translates species stat areas to species stock areas used for discard stratification based on area stratifications used in current stock assessments and quota monitoring.  Species specific area stratifications have agreed upon by a discard working group.


## Description

- **OUTPUT TABLE:** CAMS_GARFO.CAMS_STATAREA_STOCK
- **YEARS:** N/A
- **RESOLUTION:** N/A
- **DEVELOPMENT LANGUAGE:** N/A
- **CODE:** https://github.com/NOAA-Fisheries-Greater-Atlantic-Region/discaRd/tree/model_estimator/CAMS

```{r table_flow0, echo=F,fig.height=3,fig.width=8, fig.cap="Figure 1.   MAPS.CAMS_STATAREA_STOCK table lineage"}
DiagrammeR::mermaid("
  graph LR
  MAPS.CAMS_STATAREA_STOCK --> GF_and_Non_GF_Discard_Modules
        ")
```

## Data Dictionary

```{r echo = F}
comments <- ROracle::dbGetQuery(bcon, "SELECT * FROM all_col_comments WHERE owner = 'MAPS' AND table_name = 'CAMS_STATAREA_STOCK'")

comments %>%
dplyr::select(
COLUMN = COLUMN_NAME,
DESCRIPTION = COMMENTS) %>%
knitr::kable(booktabs = TRUE) %>%
kableExtra::kable_styling(bootstrap_options = c("striped", 
"hover", "condensed", "responsive"),full_width = FALSE,position = "left", fixed_thead = TRUE,
font_size = 10)
```


