---
title: "CAMS: CAMS_DISCARD_MORTALITY_STOCK"
#author: "K. Winiarski"
output:
  html_document: default
  github_document:
  pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, 
											message=FALSE, cache=FALSE,
											progress = TRUE, verbose = FALSE)


library(tidyverse)
library(odbc)
library(ROracle)
library(dplyr, warn.conflicts = FALSE)


 dw_maps <- config::get(config = "maps", file = "~/config_group.yml")
# 
 # bcon <- dbConnect(odbc::odbc(), 
 # 									DSN = dw_maps$dsn, 
 # 									UID = dw_maps$uid, 
 # 									PWD = dw_maps$pwd)

# Connect to database - move this to config file in the future - quick addition for server
  connectString <- paste(
    "(DESCRIPTION=",
    "(ADDRESS=(PROTOCOL=tcp)(HOST=", dw_maps$host, ")(PORT=", dw_maps$port, "))",
    "(CONNECT_DATA=(SERVICE_NAME=",dw_maps$svc, ")))",
    sep = ""
  )

# Connect to oracle each loop in case of timeouts
    bcon <- ROracle::dbConnect(
      drv = ROracle::Oracle(),
      username = dw_maps$uid,
      password = dw_maps$pwd,
      dbname = connectString
    )
```


## Summary
Mortality rates by area and gear from this discard estimation support table are based on values used in current stock assessments and quota monitoring.  Discard mortality rates have been agreed upon by a CAMS discard working group.

## Description

- **OUTPUT TABLE:** CAMS_DISCARD_MORTALITY_STOCK
- **YEARS:** N/A
- **RESOLUTION:** Discard mortality rates by area and gear.
- **DEVELOPMENT LANGUAGE:** N/A
- **CODE:** N/A

```{r table_flow0, echo=F,fig.height=3,fig.width=8, fig.cap="Figure 1.   CAMS_DISCARD_MORTALITY_STOCK table lineage"}
DiagrammeR::mermaid("
  graph LR
  CAMS_DISCARD_MORTALITY_STOCK --> GF_and_Non_GF_Discard_Modules
        ")
```

## Data Dictionary

```{r echo = F}
comments <- ROracle::dbGetQuery(bcon, "SELECT * FROM all_col_comments WHERE owner = 'MAPS' AND table_name = 'CAMS_DISCARD_MORTALITY_STOCK'")

comments %>%
dplyr::select(
COLUMN = COLUMN_NAME,
DESCRIPTION = COMMENTS) %>%
knitr::kable(booktabs = TRUE) %>%
kableExtra::kable_styling(bootstrap_options = c("striped", 
"hover", "condensed", "responsive"),full_width = FALSE,position = "left", fixed_thead = TRUE,
font_size = 10)
```

