---
title: "CAMS: CAMS_DISCARD_MORTALITY_STOCK"
#author: "K. Winiarski"
output:
  html_document: default
  github_document:
  pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, 
											message=FALSE, cache=FALSE,
											progress = TRUE, verbose = FALSE)


library(tidyverse)
library(odbc)
library(ROracle)
library(dplyr, warn.conflicts = FALSE)

# unlock keyring
keyring::keyring_unlock("apsd_ma")

# local run
# dw_apsd <- config::get(value = "apsd", file = "K:/R_DEV/config.yml")

# if on server..
# dw_apsd <- config::get(value = "maps", file = "~/config.yml")


# connect to MAPS

con_maps = apsdFuns::roracle_login(key_name = 'apsd_ma', key_service = 'maps')

```


## Summary
Mortality rates by area and gear from this discard estimation support table are based on values used in current stock assessments and quota monitoring.  Discard mortality rates have been agreed upon by a CAMS discard working group.

## Description

- **OUTPUT TABLE:** CAMS_DISCARD_MORTALITY_STOCK
- **YEARS:** N/A
- **RESOLUTION:** Discard mortality rates by area and gear.
- **DEVELOPMENT LANGUAGE:** N/A
- **CODE:** N/A

```{r table_flow0, echo=F,fig.height=3,fig.width=8, fig.cap="Figure 1.   CAMS_DISCARD_MORTALITY_STOCK table lineage"}
DiagrammeR::mermaid("
  graph LR
  CAMS_DISCARD_MORTALITY_STOCK --> GF_and_Non_GF_Discard_Modules
        ")
```

## Data Dictionary

```{r echo = F}
comments <- ROracle::dbGetQuery(bcon, "SELECT * FROM all_col_comments WHERE owner = 'MAPS' AND table_name = 'CAMS_DISCARD_MORTALITY_STOCK'")

comments %>%
dplyr::select(
COLUMN = COLUMN_NAME,
DESCRIPTION = COMMENTS) %>%
knitr::kable(booktabs = TRUE) %>%
kableExtra::kable_styling(bootstrap_options = c("striped", 
"hover", "condensed", "responsive"),full_width = FALSE,position = "left", fixed_thead = TRUE,
font_size = 10)
```

