---
title: "Hake Summary"
#author: "Ben Galuardi modified by K.Winiarski"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    fig_caption: true
  pdf_document: default
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning = FALSE, 
											message = FALSE, cache = FALSE,
											progress = TRUE, verbose = FALSE, comment = F
											, error = FALSE, dev = 'png', dpi = 200)

options(knitr.kable.NA = '')
```


### Background 

Exploration of discard estimates from may year module developed for hake

### Steps 

1. pull all merged trips from CAMS_OBS_CATCH

2. Set stratification variables for all trips
```{r, echo = T, eval = F}
# FULL Stratification variables

stratvars = c('SPECIES_STOCK'
              ,'CAMS_GEAR_GROUP'
							, 'MESHGROUP'
						  , 'TRIPCATEGORY'
						  , 'ACCESSAREA')
```

	- SPECIES_STOCK is taken from CAMS support table `MAPS.CAMS_STATAREA_STOCK`
	- CAMS_GEAR_GROUP is derived from a support table (`MAPS.CAMS_GEARCODE_STRATA`)
	- MESHGROUP is hardcoded for all trips according to decisions made by the mesh subgroup (see summary when available)
	- SECTID comes from a CAMS matching table (`MAPS.MATCH_MULT_SECTID`)

3. Run `discaRd`
	- there are sub-passes for year t and year t-1
	
```{r, echo = T, eval = F}
# Assumed Stratification variables

stratvars = c('SPECIES_STOCK'
              , 'CAMS_GEAR_GROUP') 
```	

4. The discaRd functions allow for an assumed rate to be calculated. This assumed rate is realtive to the stratification used in the functions. Here, the stratification is coarsened to 'SPECIES_STOCK', 'CAMS_GEAR_GROUP' and 'MESHGROUP'.

7. A transition rate is calculated between year t and year t-1. This rate determines how much, if any, information is used from previous years. 

8. Rates and `DISCARD_SOURCE` (in parentheses) are assigned for each trip according to:
 - (I) in season rate; >= 5 trips in Full Stratification
 - (T) Transition in season rate; < 5 trips in Full Stratification, year t, AND >= 5 trips in year t-1
 - (A) Assumed rate. This is the rate when there were <5 trips in season and <5 in previous season.
 - (O) Observed values used from obserevd trips; discard rate is NOT USED. 
9. CV calculations are available for (I) and (T). Obtaining a CV estimate for (A) would require a *second pass* of discaRd functions.

10. Discard pounds per trip are calculated according to

```{r echo = T, eval = F}
	mutate(coalesce(DISC_MORT_RATIO, 1)) %>%
	mutate(DISCARD = case_when(!is.na(LINK1) ~ DISC_MORT_RATIO*OBS_DISCARD
														 , is.na(LINK1) ~ DISC_MORT_RATIO*COAL_RATE*LIVE_POUNDS)
				 
			# COAL_RATE is the final discard rate used. It is 'coalesced' from the (I), (A) and (B) rates 	 
```

By assigning `SPECIES_STOCK` as a stratification variable, the computation time is reduced. Each subtrip may only occur in a single statistical area so it should never cross stock boundaries. 

Once the full table (CAMS_OBS_CATCH) is loaded, each species takes ~X seconds to process on the server.

Output tables are produced for each species. These can easily be recombined. An example table has been shared on `MAPS` and `CAMS_GARFO`.  

```sql
	MAPS.CAMS_DISCARD_EXAMPLE_CY_BARNDOORSKATE_19
	CAMS_GARFO.CAMS_DISCARD_EXAMPLE_CY_BARNDOORSKATE_19
	
	MAPS.CAMS_DISCARD_EXAMPLE_CY_CLEARNOSESKATE_19
	CAMS_GARFO.CAMS_DISCARD_EXAMPLE_CY_CLEARNOSESKATE_19
	
	MAPS.CAMS_DISCARD_EXAMPLE_CY_LITTLESKATE_19
	CAMS_GARFO.CAMS_DISCARD_EXAMPLE_CY_LITTLESKATE_19
	
	MAPS.CAMS_DISCARD_EXAMPLE_CY_ROSETTESKATE_19
	CAMS_GARFO.CAMS_DISCARD_EXAMPLE_CY_ROSETTESKATE_19
	
	MAPS.CAMS_DISCARD_EXAMPLE_CY_SMOOTHSKATE_19
	CAMS_GARFO.CAMS_DISCARD_EXAMPLE_CY_SMOOTHSKATE_19
	
	MAPS.CAMS_DISCARD_EXAMPLE_CY_WINTERSKATE_19
	CAMS_GARFO.CAMS_DISCARD_EXAMPLE_CY_WINTERSKATE_19

```

### Diagnostic Plots/Tables

```{r make diagnostic plot discards, eval = T, echo = F, fig.cap = "2019 Discard Totals by Method and Source", fig.height = 8, fig.width = 12}

library(odbc)
library(dplyr, warn.conflicts = FALSE)
# library(dbplyr)
library(ggplot2)
# library(config)
library(stringr)
library(discaRd)
library(knitr)
library(kableExtra)
options(scipen = 999)

discard <- read.csv("~/discaRd/CAMS/MODULES/MAY/Skate_comparisons.csv")

discard %>% filter(Source %in% c("Federal"))%>%
  filter(Species %in% c("Skate complex")) %>%
	ggplot()+
	geom_bar(aes(x = Method, y = Discards, fill = Source), stat = 'identity', position = 'dodge')+
	facet_wrap(~Species, scales = 'free')+
	theme_light()

```

Combined skate discard estimates are much higher for the seven skate species included in the MAPS discard mortality support table (ITIS: 564139, 160855, 564130, 564130, 564136, 564151, 564149 and 564145) compared to the 11 used for QM (160845, 160846, 160848,  160855, 564037, 564130, 564136, 564139, 564145, 564149, 564151).

Probably due to differences in gear types and discard mortality rates which I believe are just one value across skates for QM.

```{r by individual skate species, eval = T, echo = F, fig.cap = "2019 CAMS Federal Discard Totals by Skate Species", fig.height = 8, fig.width = 12}

discard <- read.csv("~/discaRd/CAMS/MODULES/MAY/Skate_comparisons.csv")

'%ni%' <- Negate("%in%")

discard %>% filter(Source %in% c("Federal")) %>%
  dplyr::select(Species, Source, Discards) %>%
		DT::datatable(caption = 'Total discards by species for state trips')

```


```{r by individual skate species state, eval = T, echo = F, fig.cap = "2019 CAMS State Discard Totals by Skate Species", fig.height = 8, fig.width = 12}

discard <- read.csv("~/discaRd/CAMS/MODULES/MAY/Skate_comparisons.csv")

'%ni%' <- Negate("%in%")

discard %>% filter(Source %in% c("State")) %>% 
  dplyr::select(Species, Source, Discards) %>%
		DT::datatable(caption = 'Total discards by species for state trips')
```

Red flag that some of those state estimates are the same between species. 


```{r make diagnostic plot kall, eval = T, echo = F, fig.cap = "2019 KALL Totals by Method and Source", fig.height = 8, fig.width = 12}

Kall <- read.csv("~/discaRd/CAMS/MODULES/MAY/May_kall.csv")

Kall %>% 
	ggplot()+
	geom_bar(aes(x = Method, y = kall, fill = Source), stat = 'identity', position = 'dodge')+
	#facet_wrap(~Species, scales = 'free')+
	theme_light()

```

Not quite sure why the QM kall estimate is so much lower.  Got it from the 2019 ACL skate complex excel file. 

```{r make diagnostic mackerel plot, eval = T, echo = F, fig.cap = "Histogram of iscard Rates by Species, Strata and Discard Source", fig.height = 8, fig.width = 12}

 dw_maps <- config::get(config = "maps", file = "~/config_group.yml")
 
# Connect to database - move this to config file in the future - quick addition for server
  connectString <- paste(
    "(DESCRIPTION=",
    "(ADDRESS=(PROTOCOL=tcp)(HOST=", dw_maps$host, ")(PORT=", dw_maps$port, "))",
    "(CONNECT_DATA=(SERVICE_NAME=",dw_maps$svc, ")))",
    sep = ""
  )

# Connect to oracle each loop in case of timeouts
    bcon <- ROracle::dbConnect(
      drv = ROracle::Oracle(),
      username = dw_maps$uid,
      password = dw_maps$pwd,
      dbname = connectString
    )

db_barndoor = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_EXAMPLE_CY_BARNDOORSKATE_19")) %>% 
	collect

db_barndoor$SPECIES <- "BARNDOOR SKATE"

db_clearnose = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_EXAMPLE_CY_CLEARNOSESKATE_19")) %>% 
	collect

db_clearnose$SPECIES <- "CLEARNOSE SKATE"

db_little = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_EXAMPLE_CY_LITTLESKATE_19")) %>% 
	collect

db_little$SPECIES <- "LITTLENOSE SKATE"

db_rosette = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_EXAMPLE_CY_ROSETTESKATE_19")) %>% 
	collect

db_rosette$SPECIES <- "ROSETTE SKATE"

db_smooth = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_EXAMPLE_CY_SMOOTHSKATE_19")) %>% 
	collect

db_smooth$SPECIES <- "SMOOTH SKATE"

db_winter = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_EXAMPLE_CY_WINTERSKATE_19")) %>% 
	collect

db_winter$SPECIES <- "WINTER SKATE"

db_final <- rbind(db_barndoor,db_clearnose,
                  db_little, db_rosette,
                  db_smooth, db_winter)

strat_barndoor <- db_barndoor %>% 	
  filter(DISCARD_SOURCE != 'O') %>%
  group_by(SPECIES, STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(
                              SPECIES,
                              STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,TRIP_TYPE)
             # ,CV) 
             #dplyr::arrange(desc(DISCARD))

strat_clearnose <- db_clearnose %>% 	
  filter(DISCARD_SOURCE != 'O') %>%
  group_by(SPECIES, STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(
                              SPECIES,
                              STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,TRIP_TYPE)

strat_little <- db_little %>% 	
  filter(DISCARD_SOURCE != 'O') %>%
  group_by(SPECIES, STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(
                              SPECIES,
                              STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,TRIP_TYPE)

strat_rosette <- db_rosette %>% 	
  filter(DISCARD_SOURCE != 'O') %>%
  group_by(SPECIES, STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(
                              SPECIES,
                              STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,TRIP_TYPE)

strat_smooth <- db_smooth %>% 	
  filter(DISCARD_SOURCE != 'O') %>%
  group_by(SPECIES, STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(
                              SPECIES,
                              STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,TRIP_TYPE)

strat_winter <- db_winter %>% 	
  filter(DISCARD_SOURCE != 'O') %>%
  group_by(SPECIES, STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(
                              SPECIES,
                              STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,TRIP_TYPE)

strat_final <- rbind(strat_barndoor, strat_clearnose,
                     strat_little, strat_rosette,
                     strat_smooth, strat_winter)
  
  
strat_final %>% 
	filter(DISCARD_SOURCE != 'O' & CAMS_DISCARD_RATE) %>% 
	dplyr::group_by(SPECIES, STRATA, DISCARD_SOURCE, TRIP_TYPE) %>% 
	#slice(1) %>% 
	ggplot()+
	geom_histogram(aes(x = CAMS_DISCARD_RATE, fill = DISCARD_SOURCE), bins=60, position = "dodge") +
	facet_wrap(~SPECIES)+
	theme_light()
```


```{r plot 1, echo = F, fig.cap = "Total Discard Estimates by Species, Stock and Discard Source" , fig.height = 8, fig.width = 12}
db_final %>% 
	# filter(DISCARD_SOURCE != 'O') %>% 
	group_by(SPECIES, SPECIES_ITIS, DISCARD_SOURCE,SPECIES_STOCK) %>% 
	dplyr::summarise(DSUM = sum(DISCARD, na.rm = T)) %>% 
	# slice(1) %>% 
	ggplot()+
	geom_bar(aes(x = SPECIES_STOCK, y = DSUM, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+
 facet_wrap(~SPECIES, scales = 'free')+
	theme_light()
```

Not suprisingly, total discards at the end of the year come from strata with an in season rate. 


```{r barndoor discard table }

db_barndoor %>% 	
  filter(DISCARD_SOURCE != 'O' & TRIP_TYPE=="FED") %>%
  group_by(STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>% top_n(10) %>%
		DT::datatable(caption = 'Discard rates by Strata for Barndoor Skate')
```



```{r clearnose discard table }

 db_clearnose %>% 	
  filter(DISCARD_SOURCE != 'O' & TRIP_TYPE=="FED") %>%
  group_by(STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>% top_n(10) %>%
		DT::datatable(caption = 'Discard rates by Strata for Clearnose Skate based on federal trips')
```

```{r little discard table }

db_little %>% 	
  filter(DISCARD_SOURCE != 'O' & TRIP_TYPE=="FED") %>%
  group_by(STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>% top_n(10) %>%
		DT::datatable(caption = 'Discard rates by Strata for Little Skate based on federal trips')
```

```{r rosette discard table }

db_rosette %>% 	
  filter(DISCARD_SOURCE != 'O' & TRIP_TYPE=="FED") %>%
  group_by(STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>% top_n(10) %>% 
		DT::datatable(caption = 'Discard rates by Strata for Rosette Skate based on federal trips')
```

```{r smooth discard table }

db_smooth %>% 	
  filter(DISCARD_SOURCE != 'O' & TRIP_TYPE=="FED") %>%
  group_by(STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>% top_n(10) %>% 
		DT::datatable(caption = 'Discard rates by Strata for Smoothnose Skate based on federal trips')
```

```{r winter discard table }

db_winter %>% 	
  filter(DISCARD_SOURCE != 'O' & TRIP_TYPE=="FED") %>%
  group_by(STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>% top_n(10) %>%
		DT::datatable(caption = 'Discard rates by Strata for Winter Skate based on federal trips')
```


### Appendix

For more information regarding stratfication variables and discard rates by species see

MAPS.CAMS_STATAREA_STOCK,
MAPS.CAMS_DISCARD_MORTALITY_STOCK,
MAPS.CAMS_GEARCODE_STRATA,
MAPS.CAMS_MASTER_GEAR,
MAPS.CAMS_OBS_CATCH,

I may need to grant selection to APSD if you do not have access to MAPS.

