---
title: "May Module Summary (Skates)"
author: "Ben Galuardi modified by K.Winiarski"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    fig_caption: true
  pdf_document: default
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning = FALSE, 
											message = FALSE, cache = FALSE,
											progress = TRUE, verbose = FALSE, comment = F
											, error = FALSE, dev = 'png', dpi = 200)

options(knitr.kable.NA = '')
```


### Background 

Exploration of discard estimates from may year module developed for skates

### Steps 

1. pull all merged trips from CAMS_OBS_CATCH

2. Set stratification variables for all trips
```{r, echo = T, eval = F}
# FULL Stratification variables

stratvars = c('SPECIES_STOCK'
              ,'CAMS_GEAR_GROUP'
							, 'MESHGROUP'
							, 'HALFOFYEAR'
							, 'REGION'
						  , 'TRIPCATEGORY'
						  , 'ACCESSAREA')
```

	- SPECIES_STOCK is taken from CAMS support table `MAPS.CAMS_STATAREA_STOCK`
	- CAMS_GEAR_GROUP is derived from a support table (`MAPS.CAMS_GEARCODE_STRATA`)
	- MESHGROUP is hardcoded for all trips according to decisions made by the mesh subgroup (see summary when available)
	- SECTID comes from a CAMS matching table (`MAPS.MATCH_MULT_SECTID`)

3. Run `discaRd`
	- there are sub-passes for year t and year t-1
	
```{r, echo = T, eval = F}
# Assumed Stratification variables

stratvars = c('SPECIES_STOCK'
              , 'CAMS_GEAR_GROUP'
              , 'MESHGROUP') 
```	

4. The discaRd functions allow for an assumed rate to be calculated. This assumed rate is realtive to the stratification used in the functions. Here, the stratification is coarsened to 'SPECIES_STOCK', 'CAMS_GEAR_GROUP' and 'MESHGROUP'.

7. A transition rate is calculated between year t and year t-1. This rate determines how much, if any, information is used from previous years. 

8. Rates and `DISCARD_SOURCE` (in parentheses) are assigned for each trip according to:
 - (I) in season rate; >= 5 trips in Full Stratification
 - (T) Transition in season rate; < 5 trips in Full Stratification, year t, AND >= 5 trips in year t-1
 - (A) Assumed rate. This is the rate when there were <5 trips in season and <5 in previous season.
 - (O) Observed values used from obserevd trips; discard rate is NOT USED. 
9. CV calculations are available for (I) and (T). Obtaining a CV estimate for (A) would require a *second pass* of discaRd functions.

10. Discard pounds per trip are calculated according to

```{r echo = T, eval = F}
	mutate(coalesce(DISC_MORT_RATIO, 1)) %>%
	mutate(DISCARD = case_when(!is.na(LINK1) ~ DISC_MORT_RATIO*OBS_DISCARD
														 , is.na(LINK1) ~ DISC_MORT_RATIO*COAL_RATE*LIVE_POUNDS)
				 
			# COAL_RATE is the final discard rate used. It is 'coalesced' from the (I), (A) and (B) rates 	 
```

By assigning `SPECIES_STOCK` as a stratification variable, the computation time is reduced. Each subtrip may only occur in a single statistical area so it should never cross stock boundaries. 

Once the full table (CAMS_OBS_CATCH) is loaded, each species takes ~X seconds to process on the server.

Output tables are produced for each species. These can easily be recombined. An example table has been shared on `MAPS` and `CAMS_GARFO`.  

```sql
	MAPS.CAMS_DISCARD_EXAMPLE_CY_BARNDOORSKATE_19
	CAMS_GARFO.CAMS_DISCARD_EXAMPLE_CY_BARNDOORSKATE_19

```

### Diagnostic Plots/Tables

```{r make diagnostic plot discards, eval = T, echo = F, fig.cap = "2019 Discard Totals by Method and Source", fig.height = 8, fig.width = 12}

library(odbc)
library(dplyr, warn.conflicts = FALSE)
library(dbplyr)
library(ggplot2)
# library(config)
library(stringr)
library(discaRd)
library(knitr)
library(kableExtra)
options(scipen = 999)

discard <- read.csv("~/discaRd/CAMS/MODULES/MAY/Skate_comparisons.csv")

discard %>% filter(Species %in% c("Skate complex")) %>%
	ggplot()+
	geom_bar(aes(x = Method, y = Discards, fill = Source), stat = 'identity', position = 'dodge')+
	facet_wrap(~Species, scales = 'free')+
	theme_light()

```




<!-- ```{r make diagnostic plot kall, eval = T, echo = F, fig.cap = "2019 KALL Totals by Method", fig.height = 8, fig.width = 12} -->

<!-- Kall <- read.csv("~/discaRd/CAMS/MODULES/MAY/May_kall.csv") -->

<!-- Kall %>%  -->
<!-- 	ggplot()+ -->
<!-- 	geom_bar(aes(x = Method, y = kall, fill = Source), stat = 'identity', position = 'dodge')+ -->
<!-- 	#facet_wrap(~Species, scales = 'free')+ -->
<!-- 	theme_light() -->

<!-- ``` -->



<!-- ```{r make diagnostic mackerel plot, eval = T, echo = F, fig.cap = "Mean Mackerel Discard Rates by Stock and Discard Source", fig.height = 8, fig.width = 12} -->

<!-- library(odbc) -->
<!-- library(dplyr, warn.conflicts = FALSE) -->
<!-- library(dbplyr) -->
<!-- library(ggplot2) -->
<!-- # library(config) -->
<!-- library(stringr) -->
<!-- library(discaRd) -->
<!-- library(knitr) -->
<!-- library(kableExtra) -->
<!-- options(scipen = 999) -->

<!--  dw_maps <- config::get(config = "maps", file = "~/config_group.yml") -->

<!-- # Connect to database - move this to config file in the future - quick addition for server -->
<!--   connectString <- paste( -->
<!--     "(DESCRIPTION=", -->
<!--     "(ADDRESS=(PROTOCOL=tcp)(HOST=", dw_maps$host, ")(PORT=", dw_maps$port, "))", -->
<!--     "(CONNECT_DATA=(SERVICE_NAME=",dw_maps$svc, ")))", -->
<!--     sep = "" -->
<!--   ) -->

<!-- # Connect to oracle each loop in case of timeouts -->
<!--     bcon <- ROracle::dbConnect( -->
<!--       drv = ROracle::Oracle(), -->
<!--       username = dw_maps$uid, -->
<!--       password = dw_maps$pwd, -->
<!--       dbname = connectString -->
<!--     ) -->

<!-- db_mack = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_EXAMPLE_CY_MACKEREL_19")) %>%  -->
<!-- 	collect() -->

<!-- db_mack %>%  -->
<!-- 	filter(DISCARD_SOURCE != 'O') %>%  -->
<!-- 	group_by(SPECIES_ITIS_EVAL, DISCARD_SOURCE, STRATA_FULL, STRATA_ASSUMED) %>%  -->
<!-- 	slice(1) %>%  -->
<!-- 	ggplot()+ -->
<!-- 	geom_bar(aes(x = SPECIES_STOCK, y = DISCARD_RATE, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+ -->
<!-- 	#facet_wrap(~COMNAME_EVAL, scales = 'free')+ -->
<!-- 	theme_light() -->
<!-- ``` -->




<!-- ```{r make diagnostic scup plot, eval = T, echo = F, fig.cap = "Mean Scup Discard Rates by Stock and Discard Source", fig.height = 8, fig.width = 12} -->
<!-- db_scup = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_EXAMPLE_CY_SCUP_19")) %>%  -->
<!-- 	collect() -->

<!-- db_scup %>%  -->
<!-- 	filter(DISCARD_SOURCE != 'O') %>%  -->
<!-- 	group_by(SPECIES_ITIS_EVAL, DISCARD_SOURCE, STRATA_FULL, STRATA_ASSUMED) %>%  -->
<!-- 	slice(1) %>%  -->
<!-- 	ggplot()+ -->
<!-- 	geom_bar(aes(x = SPECIES_STOCK, y = DISCARD_RATE, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+ -->
<!-- 	#facet_wrap(~COMNAME_EVAL, scales = 'free')+ -->
<!-- 	theme_light() -->
<!-- ``` -->


<!-- ```{r make diagnostic shortfin squid plot, eval = T, echo = F, fig.cap = "Mean Shortfin Squid Discard Rates by Stock and Discard Source", fig.height = 8, fig.width = 12} -->
<!-- db_shortfin = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_EXAMPLE_CY_SHORTFINSQUID_19")) %>%  -->
<!-- 	collect() -->

<!-- db_shortfin %>%  -->
<!-- 	filter(DISCARD_SOURCE != 'O') %>%  -->
<!-- 	group_by(SPECIES_ITIS_EVAL, DISCARD_SOURCE, STRATA_FULL, STRATA_ASSUMED) %>%  -->
<!-- 	slice(1) %>%  -->
<!-- 	ggplot()+ -->
<!-- 	geom_bar(aes(x = SPECIES_STOCK, y = DISCARD_RATE, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+ -->
<!-- 	#facet_wrap(~COMNAME_EVAL, scales = 'free')+ -->
<!-- 	theme_light() -->
<!-- ``` -->

<!-- ```{r plot 2, echo = F, fig.cap = "Total Mackerel Discard Estimates by Stock and Discard Source" , fig.height = 8, fig.width = 12} -->
<!-- db_mack %>%  -->
<!-- 	# filter(DISCARD_SOURCE != 'O') %>%  -->
<!-- 	group_by(SPECIES_ITIS_EVAL, DISCARD_SOURCE,SPECIES_STOCK) %>%  -->
<!-- 	dplyr::summarise(DSUM = sum(DISCARD, na.rm = T)) %>%  -->
<!-- 	# slice(1) %>%  -->
<!-- 	ggplot()+ -->
<!-- 	geom_bar(aes(x = SPECIES_STOCK, y = DSUM, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+ -->
<!-- #	facet_wrap(~COMNAME_EVAL, scales = 'free')+ -->
<!-- 	theme_light() -->
<!-- ``` -->



<!-- ```{r plot 3, echo = F, fig.cap = "Total Scup Discard Estimates by Stock and Discard Source" , fig.height = 8, fig.width = 12} -->
<!-- db_scup %>%  -->
<!-- 	# filter(DISCARD_SOURCE != 'O') %>%  -->
<!-- 	group_by(SPECIES_ITIS_EVAL, DISCARD_SOURCE,SPECIES_STOCK) %>%  -->
<!-- 	dplyr::summarise(DSUM = sum(DISCARD, na.rm = T)) %>%  -->
<!-- 	# slice(1) %>%  -->
<!-- 	ggplot()+ -->
<!-- 	geom_bar(aes(x = SPECIES_STOCK, y = DSUM, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+ -->
<!-- #	facet_wrap(~COMNAME_EVAL, scales = 'free')+ -->
<!-- 	theme_light() -->

<!-- ``` -->

<!-- ```{r plot 4, echo = F, fig.cap = "Total Shortfin Squid Discard Estimates by Stock and Discard Source" , fig.height = 8, fig.width = 12} -->
<!-- db_shortfin %>%  -->
<!-- 	# filter(DISCARD_SOURCE != 'O') %>%  -->
<!-- 	group_by(SPECIES_ITIS_EVAL, DISCARD_SOURCE,SPECIES_STOCK) %>%  -->
<!-- 	dplyr::summarise(DSUM = sum(DISCARD, na.rm = T)) %>%  -->
<!-- 	# slice(1) %>%  -->
<!-- 	ggplot()+ -->
<!-- 	geom_bar(aes(x = SPECIES_STOCK, y = DSUM, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+ -->
<!-- #	facet_wrap(~COMNAME_EVAL, scales = 'free')+ -->
<!-- 	theme_light() -->

<!-- ``` -->

<!-- # ```{r table 2} -->
<!-- # #Bring in strata for mackerel for DMIS 2019. -->

<!-- # #Bring in strata for scup for DMIS 2019. -->

<!-- # ``` -->


<!-- # ```{r discard rates by strata mackerel example} -->

<!-- # db_mack %>%  -->
<!-- # 	group_by(DISCARD_SOURCE -->
<!-- # 					 , STRATA_FULL -->
<!-- # 					 , STRATA_ASSUMED) %>%  -->
<!-- # 	dplyr::summarise(DISCARD_RATE = max(DISCARD_RATE) -->
<!-- # 									 , KALL = sum(SUBTRIP_KALL, na.rm = T)) %>%  -->
<!-- # 	knitr::kable(format.args = list(big.mark = ",")) -->

<!-- # ``` -->

<!-- # ```{r discard rates by strata scup example} -->
<!-- # db_scup %>%  -->
<!-- # 	group_by(DISCARD_SOURCE -->
<!-- # 					 , STRATA_FULL -->
<!-- # 					 , STRATA_ASSUMED) %>%  -->
<!-- # 	dplyr::summarise(DISCARD_RATE = max(DISCARD_RATE) -->
<!-- # 									 , KALL = sum(SUBTRIP_KALL, na.rm = T)) %>%  -->
<!-- # 	knitr::kable(format.args = list(big.mark = ",")) -->

<!-- # ``` -->

### Summary

1. Combined skate discard estimates are much higher for the seven skate species included in the MAPS discard mortality support table (ITIS: 564139, 160855, 564130, 564130, 564136, 564151, 564149 and 564145) compared to the 17 used for QM (160845, 160846, 160848,  160855, 564037, 564130, 564136, 564139, 564145, 564149, 564151).


### Current Issues with May Year Module (skates)

1. KALL is low for QM.

2. Some trips are not getting a final discard rate (mostly state trips but some federal subtrips). 

3. Having issues writing some species tables to MAPS from R.


### Next Steps for May Year Module (skates)

1. Need to make sure I am including relevant QM nuances in calculation of  skate discards.

4. Add summary to CAMS documentation once complete.

