---
title: "CAMS:CAMS_DISCARDS_ALL_YEARS"
#author: "Ben Galuardi"
output:
  html_document: default
  github_document:
  pdf_document: default
always_allow_html: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(odbc)
library(ROracle)
library(dplyr, warn.conflicts = FALSE)

# # unlock keyring
# keyring::keyring_unlock("apsd_ma")
# 
# # local run
# # dw_apsd <- config::get(value = "apsd", file = "K:/R_DEV/config.yml")
# 
# # if on server..
# # dw_apsd <- config::get(value = "maps", file = "~/config.yml")
# 
# 
# # connect to MAPS
# 
# con_maps = apsdFuns::roracle_login(key_name = 'apsd_ma', key_service = 'maps')


```

```{r, include = FALSE, echo = FALSE}
Sys.setenv(TZ = "America/New_York")
Sys.setenv(ORA_SDTZ = "America/New_York")
  
if(file.exists(here::here("vignettes/articles", "pw.csv"))) {
  pw_df <- readr::read_csv(here::here("vignettes/articles", "pw.csv"), col_types = "c")
  pw <- pw_df$pw
} else {
  pw <- rstudioapi::askForPassword("Enter keyring password")
}

keyring::keyring_unlock("apsd_ma", password = pw) # will prompt for password
con_maps <- apsdFuns::roracle_login(key_name = 'apsd_ma', key_service = 'maps')
```

## Summary
This table provides discard estimates by subtrip for both GF and non GF species. Table output is from the GF and Non GF discard modules available in the R package discaRd.

## Description

- **OUTPUT TABLE:** CAMS_DISCARDS_ALL_YEARS
- **YEARS:** 200X+
- **RESOLUTION:** Species discard estimate by VTRSERNO (subtrip)
- **DEVELOPMENT LANGUAGE:** N/A
- **CODE:** https://github.com/NOAA-Fisheries-Greater-Atlantic-Region/discaRd/tree/model_estimator/CAMS


## Data Sources
CAMS_OBS_CATCH
GF_EM_DISCARDS_ARCH
GF_EM_DELTA_VTR_DISCARDS
CAMS_DISCARD_MORTALITY_STOCK
CAMS_OBS_CATCH
CAMS_GEARCODE_STRATA
CAMS_STATAREA_STOCK

```{r table_flow_1, echo=F,fig.height=3,fig.width=8, fig.cap="Figure 1.   CAMS_DISCARD_ALL_YEARS table lineage for GF species"}
DiagrammeR::mermaid("
  graph LR
  CAMS_OBS_CATCH --> GF_Discard_Module
  CAMS_DISCARD_MORTALITY_STOCK --> GF_Discard_Module
  CAMS_GEARCODE_STRATA --> GF_Discard_Module
  CAMS_STATAREA_STOCK --> GF_Discard_Module
  CAMS_GF_EM_DISCARDS_ARCH --> GF_Discard_Module
  CAMS_GF_EM_DELTA_VTR_DISCARDS --> GF_Discard_Module
  GF_Discard_Module --> CAMS_DISCARDS_ALL_YEARS
        ")
```
## Approach
GF module:

1. Pull all merged trips from `CAMS_OBS_CATCH`.

2. Filter to only Groundfish trips (Use `GF` column)

3. Set stratification variables for groundfish trips

```{r, echo = T, eval = F}
# FULL Stratification variables

stratvars = c( 'SPECIES_ESTIMATION_REGION'
              , 'CAMS_GEAR_GROUP'
              , 'MESHGROUP'
              , 'SECTID'
              , 'EM'
              , "REDFISH_EXEMPTION"
              , "SNE_SMALLMESH_EXEMPTION"
              , "XLRG_GILLNET_EXEMPTION")
```
By Species:
	- SPECIES_ESTIMATION_REGION is taken from CAMS support table `CAMS_STATAREA_STOCK`
	- CAMS_GEAR_GROUP is derived from a support table (`CAMS_GEARCODE_STRATA`)
	- MESHGROUP is hardcoded for all trips according to decisions made by the mesh subgroup (see summary when available)
	- SECTID comes from a CAMS matching table (`MATCH_MULT_SECTID`)
	- EFP and Exemptions are built in to `CAMS_LANDINGS`, the precursor for `CAMS_OBS_CATCH`
	
4. Perform *first pass* of `discaRd`
	- there are two sub-passes for year t and year t-1
5. Perform *second pass* of `discaRd` with discard rates rolled up for all Sectors
	- Common Pool is distinguished from the rest of Sectors
	- Simplified stratification is used:
	
```{r, echo = T, eval = F}
# Assumed Stratification variables

stratvars = c('SPECIES_ESTIMATION_REGION'
              ,'CAMS_GEAR_GROUP'
              ,'MESHGROUP'
							,'SECTOR_TYPE') 
```	

6. The discaRd functions allow for an assumed rate to be calculated. This assumed rate is relative to the stratification used in the functions. Here, we utilize this feature to generate a broad stock rate. the stratification here is simply `SPECIES_ESTIMATION_REGION`

7. For each *pass*, a transition rate is calculated between year t and year t-1. This rate determines how much, if any, information is used from previous years. 

8. The two *passes* are joined in a hierarchical manner. Rates and `DISCARD_SOURCE` (in parentheses) are assigned for each trip according to:
 - (I) in season rate; >= 5 trips in Full Stratification
 - (T) Transition in season rate; < 5 trips in Full Stratification, year t, AND >= 5 trips in year t-1
 - (A) Assumed rate. This is the *second pass* rate when there were >=5 trips in season
 - (B) Broad stock rate is used when other criteria are not met. 
 - (O) Observed values used from observed trips; discard rate is NOT USED. 
 - (VTR) VTR estimates from EM trips prior to implementation of the audit model during GF FY21.
 - (Delta) Delta estimates from EM trips after implementation of the audit model.
 - (EM) EM estimates from EM trips where EM was reviewed. 
 
9. CV calculations are available for (I), (T), and (A). Obtaining a CV estimate for (B) would require a *third pass* of discaRd functions. (O) rates are not used and final discard values are not estimated. 

10. Discard pounds per trip are calculated according to

```{r echo = T, eval = F}
	mutate(coalesce(DISC_MORT_RATIO, 1)) %>%
	mutate(DISCARD = case_when(!is.na(LINK1) ~ DISC_MORT_RATIO*OBS_DISCARD
														 , is.na(LINK1) ~ DISC_MORT_RATIO*COAL_RATE*LIVE_POUNDS)
				 
```

### Steps for estimation of groundfish discard on non-Groundfish Trips

1. same as above

2. same as above

3. Set stratification variables

```{r, echo = T, eval = F}
# Full (First Pass) Stratification variables

stratvars = c('SPECIES_ESTIMATION_REGION'
              ,'CAMS_GEAR_GROUP'
							, 'MESHGROUP'
						  , 'TRIPCATEGORY'
						  , 'ACCESSAREA')
```

	- SPECIES_ESTIMATION_REGION is taken from CAMS support table `CAMS_STATAREA_STOCK`
	- CAMS_GEAR_GROUP is derived from a support table (`CAMS_GEARCODE_STRATA`)
	- MESHGROUP is hardcoded for all trips according to decisions made by the mesh subgroup (see summary when available)
	- TRIPCATEGORY is specific to scallop trips and indicates Limited Access or General Category Fleets
	- ACCESS AREA is specific to scallop trips and indicates Scallop Access Area

4. Run first pass of `discaRd` for full startififcation
	- there are sub-passes for year t and year t-1
	
5. Run a second pass using a coarsened stratification : 'SPECIES_ESTIMATION_REGION', 'CAMS_GEAR_GROUP' and 'MESHGROUP'.

	
```{r, echo = T, eval = F}
# Second Pass Stratification variables

stratvars = c('SPECIES_ESTIMATION_REGION'
              ,'CAMS_GEAR_GROUP'
              ,'MESHGROUP')

```	


6. Run a third pass using Stock and Gear only
*Note: This pass uses data from the focal and previous year. 
```{r, echo = T, eval = F}
# Third PAss Stratification variables

stratvars = c('SPECIES_ESTIMATION_REGION'
              ,'CAMS_GEAR_GROUP')

```	

7. Trips/Gear strata that are unobserved after these steps are not assigned rates as no observer coverage exists for these gear types (e.g. Menhaden purse seine)

8. Rates and `DISCARD_SOURCE` (in parentheses) are assigned for each trip according to:
 - (I) in season rate; >= 5 trips in Full Stratification
 - (T) Transition in season rate; < 5 trips in Full Stratification, year t, AND >= 5 trips in year t-1
 - (GM) Broad stock/gear/mesh size rate. This is the rate when there were <5 trips in season and <5 in previous season.
 - (G) Broad stock/gear rate. <5 trips in season and <5 trips in previous season
 - (N) No coverage in focal season and previous season; discard rate is not possible to estimate.
 - (O) Observed values used from observed rate trips; discard rate is NOT USED. 

9. CV calculations are available for (I), (T), (GM) and (G). 

10. Discard pounds per trip are calculated according to

```{r echo = T, eval = F}
	mutate(DISC_MORT_RATIO = coalesce(DISC_MORT_RATIO, 1)) %>%
	mutate(DISCARD = case_when(!is.na(LINK1) ~ DISC_MORT_RATIO*OBS_DISCARD
														 , is.na(LINK1) ~ DISC_MORT_RATIO*COAL_RATE*LIVE_POUNDS)
				 
```

```{r table_flow_2, echo=F,fig.height=3,fig.width=8, fig.cap="Figure 2.   CAMS_DISCARD_ALL_YEARS table lineage"}
DiagrammeR::mermaid("
  graph LR
  CAMS_OBS_CATCH --> Non_GF_Discard_Module
  CAMS_DISCARD_MORTALITY_STOCK --> Non_GF_Discard_Module
  CAMS_GEARCODE_STRATA --> Non_GF_Discard_Module
  CAMS_STATAREA_STOCK --> Non_GF_Discard_Module
  Non_GF_Discard_Module --> CAMS_DISCARDS_ALL_YEARS
        ")
```

## Approach
Non GF module:

1. pull all merged trips from CAMS_OBS_CATCH

2. Set stratification variables for all trips

```{r, echo = T, eval = F}
# FULL Stratification variables

stratvars = c('SPECIES_ESTIMATION_REGION'
              ,'CAMS_GEAR_GROUP'
							, 'MESHGROUP'
						  , 'TRIPCATEGORY'
						  , 'ACCESSAREA')
```
By Species:
	- SPECIES_ESTIMATION_REGION is taken from CAMS support table `CAMS_STATAREA_STOCK`
	- CAMS_GEAR_GROUP is derived from a support table (`CAMS_GEARCODE_STRATA`)
	- MESHGROUP is hardcoded for all trips according to decisions made by the mesh subgroup (see summary when available)
	- SECTID comes from a CAMS matching table (`MATCH_MULT_SECTID`)

3. Run `discaRd`
	- there are sub-passes for year t and year t-1
	
```{r, echo = T, eval = F}
# Assumed Stratification variables

stratvars = c('SPECIES_ESTIMATION_REGION'
              , 'CAMS_GEAR_GROUP'
              ,'MESHGROUP')

```	

4. The discaRd functions allow for an assumed rate to be calculated. This assumed rate is relative to the stratification used in the functions. Here, the stratification is coarsened to 'SPECIES_ESTIMATION_REGION', 'CAMS_GEAR_GROUP' and 'MESHGROUP'.

7. A transition rate is calculated between year t and year t-1. This rate determines how much, if any, information is used from previous years. 

8.  A broad stock/gear stratification is applied to subtrips that do not meet the requirements for the full or assumed stratification.  This broad stock/gear stratification utilizes observed coverage from both the focal year and the previous year as it includes gear types with low observer coverage.

8. Rates and `DISCARD_SOURCE` (in parentheses) are assigned for each trip according to:
 - (I) in season rate; >= 5 trips in Full Stratification
 - (T) Transition in season rate; < 5 trips in Full Stratification, year t, AND >= 5 trips in year t-1
 - (GM) Broad stock/gear/mesh size rate. This is the rate when there were <5 trips in season and <5 in previous season.
 - (G) Broad stock/gear rate. <5 trips in season and <5 trips in previous season 
 - (N) No coverage in focal season and previous season; discard rate is not possible to estimate.
 - (O) Observed values used from observed rate trips; discard rate is NOT USED. 

9. CV calculations are available for (I), (T), (GM) and (G). 

10. Discard pounds per trip are calculated according to

```{r echo = T, eval = F}
	mutate(DISC_MORT_RATIO = coalesce(DISC_MORT_RATIO, 1)) %>%
	mutate(DISCARD = case_when(!is.na(LINK1) ~ DISC_MORT_RATIO*OBS_DISCARD
														 , is.na(LINK1) ~ DISC_MORT_RATIO*COAL_RATE*LIVE_POUNDS)
				 
```


## Data Dictionary
```{r echo = F}
comments <- ROracle::dbGetQuery(con_maps, "SELECT * FROM all_col_comments WHERE owner = 'MAPS' AND table_name = 'CAMS_DISCARD_MONKFISH_19'")

comments %>%
dplyr::select(
COLUMN = COLUMN_NAME,
DESCRIPTION = COMMENTS) %>%
knitr::kable(booktabs = TRUE) %>%
kableExtra::kable_styling(bootstrap_options = c("striped", 
"hover", "condensed", "responsive"),full_width = FALSE,position = "left", fixed_thead = TRUE,
font_size = 10)
```



	
