---
title: "discaRd example using eflalo"
author: "Benjamin Galuardi"
date: 'Documented on: `r Sys.Date()`'
output: rmarkdown::html_vignette
vignette: '
%\VignetteIndexEntry{SBRM Examples}
%\VignetteEngine{knitr::rmarkdown}
%\usepackage[utf8]{inputenc}'
---

## Load trip data (dm)
We include European fisheries data (*eflalo*) to demonstrate the major capabilities of the discaRd package. This original data is courtesy of ICES ([www.ices.dk](http://ices.dk/Pages/default.aspx))and Niels Hintzen, and can be found in the [vmstools package for R](http://nielshintzen.github.io/vmstools/)


```{r, echo = T, warning = F, message = F, eval = T}
library(discaRd)
data(eflalo, package = 'discaRd')
```

## Using the *make.obs.flag.dat* function, any level of observer covereage, as well as other discard species assumptions, may be made by changing this criteria. 

This function modifies the eflalo dataset to include columns necessary to work with discaRd functions. It also assigns a percentage of trips as observed (e.g. 10%).  


```{r make_obs_flag, echo = T, warning = F, message = F, eval = T}
dm = make.obs.flag.dat(eflalo, obs_level = .1)
str(dm)
```

## Set up data for Cochran estimator

We assume an arbitrary bycatch species, European seabass (_Dicentrarchus labrax_). All species names are based on [FAO species codes](https://datacollection.jrc.ec.europa.eu/wordef/fao)

```{r setup_data, echo = T, warning = F, message = F, eval = T}
bspec = 'LE_KG_BSS' # European seabass (FAO code BSS)

# Observed subset
bdat = ddply(dm[dm$OBSFLAG==1,], 'STRATA' ,function(x) get.bydat(x, aggfact = 'DOCID',load = F, bspec = bspec, catch_disp = 1))

# All trips
ddat = dm[,c('FY','STRATA','DOCID','DATE_TRIP','fday','yday','NESPP3','LIVE_POUNDS')]
```

##Run the Cochran estimator

Follow the same process as the GARFO peer review (see [here](https://www.greateratlantic.fisheries.noaa.gov/aps/discard/review/index.html))
```{r run_cochran, echo = T, warning = F, message = F}
# setup the unique strata in full dataset
strata_complete =  unique(c(bdat$STRATA, ddat$STRATA))

# set a year of interest
focal_year = 1801

bydat_focal = subset(bdat, FY == focal_year)
bydat_prev = subset(bdat, FY == focal_year - 1)
trips_focal = subset(ddat, FY == focal_year)
trips_prev = subset(ddat, FY == focal_year - 1)

# first day of the fishing year with a commerical trip
minday = min(c(unique(subset(bdat, FY==focal_year)$fday)), unique(subset(ddat, FY==focal_year)$fday))

dest_y1 <- get.cochran.ss.by.strat(bydat_focal, trips_focal, targCV =.3, strata_name = "STRATA", strata_complete = strata_complete)

# previous year
dest_y0 <- get.cochran.ss.by.strat(bydat_prev, trips_prev, targCV =.3, strata_name = "STRATA", strata_complete = strata_complete)

# compare the two years total discard rate
data.frame(r = c(dest_y0$rTOT, dest_y1$rTOT)
, CVTOT = c(dest_y0$CVTOT, dest_y1$CVTOT)
, row.names = c('y0','y1')
)

```

## Run with a transition rate
```{r use_trans_rate, echo = T, results='asis', warning = F, message = F, eval = T}
# Use the Cochran Trans function
	dest <- cochran.trans.calc(bydat_focal = bydat_focal, trips_focal = trips_focal, bydat_prev = bydat_prev, trips_prev = trips_prev, CV_target =.3, strata_name = "STRATA", strata_complete = strata_complete, time_span = c(minday, 365))

str(dest$D)
```


## Plot the total discard
```{r plot_discard, echo = F, eval = T, message = F, warning = F}
	# get cumulative discard at each day...
	cdest = colSums(dest[[1]], na.rm=T)
	cdest[is.nan(cdest)] = 0
	cdest[is.na(cdest)] = 0
	
# add catch cap number here.. arbitrary 10% more than the discard estimate.. 
cc =  max(cdest)*1.1  

# dates of the fishing year
fdates = seq(mdy(paste0('01-01-', focal_year)), mdy(paste0('12-31-',focal_year)), 'day')

plot(fdates[as.numeric(names(cdest))], cdest, lwd =2 , col = 4, typ='l', xlab='Fishing Day', ylab='Discard', ylim = c(0, cc*1.2))

abline(h = cc, lwd = 2, lty=2, col = 2)
grid()
```

## Bootstrap example: 100 iterations in parallel (NOT RUN)
```{r boot, echo = T, eval = F, message = F, warning = F}
	nboot = 100
	ncores = detectCores()
	cl = makeCluster(ncores)
	registerDoParallel(cl, cores = ncores)
	
	(t1 =Sys.time())
	print(paste0('Bootstrapping ', nboot, ' times using ', ncores, ' cores'))
	bout.list = foreach(1:nboot) %dopar% {
	library(discaRd)
	discaRd::bootr.strat(bdat = bdat, ddat = ddat, focal_year = focal_year, strata_name = 'STRATA', strata_complete = strata_complete, time_inter = 7, trans_method = "ntrips", time_span = c(minday, 365))
	}
	(t2 = Sys.time()-t1)
	
	# Stop cluster
	stopCluster(cl)
```

## Plot bootstrapped results (NOT RUN)
```{r plot_boot, eval = F, echo=F, fig.height=7, fig.width=8, message=F, warning=F}
# dates of the fishing year
	fdates = seq(mdy(paste0('01-01-', focal_year)), mdy(paste0('12-31-',focal_year+1)), 'day')

# use subset columns
idx = as.numeric(colnames(bout.list[[1]]$D))

#------------------------------------------#
# plot by quantile
#------------------------------------------#

bdf = ldply(bout.list, function(x) colSums(x$D, na.rm = T))

bdf = t(apply(bdf, 2, function(x) quantile(x,  probs = c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=T)))


plot(fdates[as.numeric(names(cdest))], cdest, lwd =2 , col = 4, typ='l', xlab='Fishing Day', ylab='Discard', ylim = c(0, cc*1.2))

matplot(fdates[idx], bdf, typ='l', col = c(1,2,3,2,1), lty = c(3,2,1,2,3), ylab = 'Discard', xlab = 'Fishing Day', ylim = c(0, max(bdf)*1.05), add=T)
# lines(fdates[1:365], cdest, lwd =2 , col = 4, add=T)

grid()

legend('bottomright', c('Current estimate','Median','50%','95%'), lty = c(1,1,2,3), col = c(4,3,2,1), lwd = c(2,1,1,1), bg = 'white')

# catch cap
abline(h = cc, lwd = 2, lty=2, col = 2)
text(fdates[100], cc*1.05, paste0(focal_year, ' Catch Cap'))    # adjust the annotation location

title(paste0('European seabass', '\n', 'FY ', focal_year, ': 5 trip based Transition Rate'))

```




