---
title: "CAMS discaRd Update"
author: 
  - Ben Galuardi & Dan Linden
  - ^[APSD]
date: "`r lubridate::today()`"  
output:
  beamer_presentation:
    highlight: haddock
    includes:
    keep_tex: yes
    theme: Singapore
editor_options:
  chunk_output_type: inline
---


## Discard Rate Current Method Summary (J. Michael Lanning summary)

1. Rates determine by observer reported values (gear, area, etc)
2. Incomplete observed trips have missing 'hauls' prorated by observed information from that trip
3. Trips with observer get reported/calculated observed discards   of that specific trip
4. Unobserved trips get discards from the rate calculated from 1)
5. QM is only interested in the summary total of discards for each trip, not subtrips. Often the interested number is a summary of trips, ie. the herring total of bycatch for an area/season or a sector's season's total of GB Cod.
6. Other others are driven by regs. Here I would place transition rates and EM methods.

<!-- I recommend separating out issues such as mismatching gear and area for the future QA data system.  -->

<!-- And of course, any change management must be worked through proper and transparent interaction with all end users and clients including council and  SFD. -->

## discaRd Base and Support tables

```{r table_flow0, echo = F, fig.height = 8, fig.width = 10, fig.cap = "Base tables (rectangle), Intermediary (circle), and Support tables (Oval)", warning = F}
library(DiagrammeR)

grViz("
digraph boxes_and_circles {

  # a 'graph' statement
  graph [overlap = false, fontsize = 32, fillcolor = red]

  # several 'node' statements
  node [shape = box,
        fontname = Helvetica
        , fontsize = 32
        fontcolor = 'darkgreen', 
        penwidth = 2]
  'CAMS 
  Appt'; 
  
  'CAMS 
  trip attr';
  
  'NEFOP';
  
  'ASM';

 node [shape = oval,
        fontname = Helvetica
        , fillcolor = 'Cyan']
  'CAMS
  STATAREA
  STOCK'; 
  
  'CAMS
  GEARCODE
  STRATA'; 
  
  'CAMS
  DISCARD
  MORTALITY
  STOCK';

  node [shape = circle,
        fixedsize = false,
        width = 2,  fontsize = 32] // sets as circles
  'CAMS
  OBS
  CATCH'; 
  
  'OBS 
   by YEAR' ; 
   
   'CAMS
   CATCH
   TA
   MOCK'; 
   
   'OBS
   CAMS
   PRORATE';
  
  node [shape = diamond,
        fixedsize = false,
        width = 2,  fontsize = 40, fontweight = 'bold'] // 
  'R';
  

  # several 'edge' statements

  'NEFOP'->'OBS 
   by YEAR'

  'ASM'->'OBS 
   by YEAR'

  'OBS 
   by YEAR'->'OBS
   CAMS
   PRORATE'
  
  'CAMS 
  Appt' -> 'CAMS
   CATCH
   TA
   MOCK'
  'CAMS 
  trip attr' ->'CAMS
   CATCH
   TA
   MOCK'
  
  'CAMS
   CATCH
   TA
   MOCK'->  'CAMS
  OBS
  CATCH'
  
'OBS
   CAMS
   PRORATE'->  'CAMS
  OBS
  CATCH'
  
  'CAMS
  OBS
  CATCH'->'R'
  
  'CAMS
  STATAREA
  STOCK'-> 'R'
  
  'CAMS
  GEARCODE
  STRATA'->'R'
  
  'CAMS
  DISCARD
  MORTALITY
  STOCK'->'R'
  
 
}
")

```


<!-- ## Tables created and steps to date -->

<!-- **Rates determine by observer reported values (gear, area, etc)** -->
<!-- Observed discards -->

<!-- `make_obdbs_table_cams_v2.sql` -->

<!-- created:  -->

<!-- * `apsd.bg_obdbs_cams_mock2018` -->
<!-- * `apsd.bg_obdbs_cams_mock2019` -->
<!-- * `apsd.bg_obdbs_cams_mock2020` -->

## Prorated discards
**Incomplete observed trips have missing 'hauls' prorated by observed information from that trip**

Prorate observed discards on unobserved hauls within a subtrip.
This is done by applying a ratio of kept all on the entire trip to kept all on the unobserved hauls only

$$d_{total} = d_{observedhauls}*(1+KALL_{unobserved hauls}/KALL_{subtrip})$$

<!-- `make_obdbs_prorate.sql` -->

<!-- created: -->

<!-- * `apsd.obs_cams_prorate`   -->
<!-- 	+ this table was made using  `apsd.bg_obdbs_cams_mock2018` and `apsd.bg_obdbs_cams_mock2019` -->

<!-- ## Use prorated observed discard values  -->
<!-- **Trips with observer get reported/calculated observed discards of that specific trip** -->

<!-- Match observed hauls to subtrips -->

<!-- `explore_link3_mesh_match.sql` -->

<!-- This step matches on `LINK1` -->
<!-- `AREA`, `GEAR` and `MESHGROUP` (sm, lg, xlg). This is a hard match and will go awry if there is a mismatch in the data. -->

<!-- ## Use prorated observed discard values (cont.) -->
<!-- tables created:  -->

<!-- * `apsd.bg_cams_catch_mock` -->
<!-- 	+ follows the steps layed out for mid-Atlantic discard estimation. `Gear`, `mesh`, `region`, `half of year` CASE statements should be replaced at some point with table driven code.   -->
<!-- 	+ Utilizes the current apportionment table: `apsd.cams_apport_20201230` -->
<!-- * `apsd.bg_obs_cams_tmp1` -->
<!-- 	+ links to `dmis.d_match_obs_link` and `apsd.bg_cams_catch_mock`; **multiple** subtrips only.  -->
<!-- <!-- * to date (`r lubridate::today()`), this is only using trips that have **multiple** subtrips. These are the only cases where the pro-ration step matters.  --> 
<!-- *  `apsd.bg_obs_cams_tmp2` is used in the `squid example` and include all trips.  -->

<!-- **Unobserved trips get discards from the rate calculated from 1)** -->


## R Process

`discaRd` R package built for 2016 Discard Estimation Peer Review

\includegraphics[width = .25\textwidth, height = .25\textheight]{discaRd.png}

New functions for CAMS:

<!-- - `get_obs_disc_vals`  -->
- `make_assumed_rate` Calculates 'fallback rate' using a subset of `STRATA` variables
- `make_bdat_focal` Constructs data frame of observed trip data for species of interest
- `run_discard` Runs these functions in conjucntion with `discaRd`

## Running it 

- refresh Oracle tables?
- define species and stock (if applicable)
	+ generates SQL
- import to R
	+ apply CAMS_GEAR_GROUP according to SPECIES
	+ apply STOCK STAT AREA according to SPECIES and stock (if needed)
	+ join discard mortality by species/stock/CAMS_GEAR_GROUP
- `run_discard`
	+ STRATA is assigned dynamically by using elements of the imported data
	+ If using `transition rates`, two time periods are defined
	+ Assumed (fallback rates) are defined as a subset of STRATA
- Apply Discard Mortality 
	

<!-- ## output -->

<!-- \includegraphics[width = \textwidth]{dest_strata_squid_ex.jpg} -->
<!-- \vspace{2cm} -->
<!-- \includegraphics[width = \textwidth]{squid_ex_table.jpg} -->

<!-- ## compare to ACL summary -->
<!-- \includegraphics[width = \textwidth]{CAMS_discaRd_ACL_comparison.jpg} -->

## TO DO
- utilize support tables
	+ `CAMS_GEAR_GROUP` **DONE**
	+ `STAT_AREAS` **DONE**
	+ `CAMS_DISCARD_MORTALITY_STOCK` **in process**
- add `SECTOR` for multispecies (see above) **DONE**
- Time periods **Determined by STOCK/SPECIES**
	+ Species with the same time period, e.g. Calendar year, can be imported at once. 
- Assumed (fallback) rate criteria: how simplified must this be?
- implement transitions (if using fixed time period) **DONE**
- Deal with Exemptions
- Incorporate stratification for EM trips
- refine exact operational process*

***Will likely be based on modules that run common sets of species (e.g. common `CAMS_GEAR_GROUP` and stock definition)**

## Modules

<!-- We envision CAMS discards to run via several modules that are specific to the FMP and purpose. They will include: -->
<!-- \begin{small} -->

- Quota Monitoring
	+ Squid/Mackerel/Butterfish (Calendar year)
		++ This may encompass any UNIT stock with calendar year
	+ Groundfish (May year)
	+ Monkfish (May year)
	+ Yellowtail/Windowpane in scallop fishery (April year)
	+ Skates (?)
	+ Small mesh species (hakes)
	+ Dogfish (May year)
	+ ??
- SBRM
	+ 300 species on SBRM year (Calendar?)
-Stock Assessments
	+ typically run on calendar years for all species

<!-- \end{small} -->


```{r}

```
