---
title: "CAMS: Prorated Observer Table"
author: "Ben Galuardi"
output:
  html_document: default
  pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary


## Description

- **OUTPUT TABLE:** APSD.BG_CAMS_OBS_CATCH
- **YEARS:** 2018-2020
- **RESOLUTION:** VTRSERNO (subtrip)~LINK1
- **DEVELOPMENT LANGUAGE:** SQL
- **CODE:** https://github.com/NOAA-Fisheries-Greater-Atlantic-Region/discaRd/tree/model_estimator/CAMS


## Data Sources
- CAMS Apportionment and Trip attributes (GARFO)
- CAMS prorated observer data
	+ NEFOP (NEFSC)
	+ ASM (NEFSC)

```{r table_flow0, echo=F,fig.height=3,fig.width=8, fig.cap="Figure 1.   APSD.BG_CAMS_OBS_CATCH table lineage"}
DiagrammeR::mermaid("
  graph LR
  apsd.cams_trip_attribute --> APSD.BG_CAMS_CATCH
  apsd.cfders_vtr_apportionment --> APSD.BG_CAMS_CATCH
  NEFOP --> APSD.OBS_CAMS_PRORATE
  ASM --> APSD.OBS_CAMS_PRORATE
  APSD.BG_CAMS_CATCH --> APSD.BG_CAMS_OBS_CATCH
  APSD.OBS_CAMS_PRORATE --> APSD.BG_CAMS_OBS_CATCH

        ")
```
## Approach

The use of a combined catch and observation table allows for a single source table to be used in discard estimation. Previous methods took a two table approach, where catch information and observer records were stratified independently, and then matched to calculate discard (e.g., $D = K*d/k$). This approach had the possibility of mismatches between observed strata and trip strata. In reality, this cannot occur. We therefore do trip by trip matching, using `gear`, `mesh`, `statistical area`, and `LINK1`, to match observer recorded species discards with commercial trip activity. 

The primary driver for this approach was to use the trip recorded metrics as the stratification source. This reduces the possibility of mismatches and removes much of the hard-coding that has been used to date. Upfront matching also allows observed discards to easily be used as the official discard for a particular trip. Furthermore, we recognize that data errors either from the catch, or observer data, will result in a non-match. This likely reduces the total pool of observed trips that are being used, but we feel the benefits of using outweigh a reduced sample size. Quality control of these data are outside the purview of the CAMS project. Last, we only use observed trips (`LINK1`) where valid `LINK3` (hauls) occurred to alleviate issues of multiple `LINK1` records for a single subtrip. 

IMPORTANT!: The table itself is a hybrid. For trips that were not observed, there will be a single row with all trip metrics and a total `KALL` per subtrip. When a trip was observed, there are multiple rows where the trip metric information is repeated, and each row shows species, discarded amount, and other observer recorded information for each row. Total `KALL` CANNOT be calculated without filtering rows by `LINK1` to indicate an observed trip or not. These steps are outlined in subsequent R modules used to run disacRd. 

## Data Dictionary

```{r echo = F}

library(readxl)
library(knitr)

dat = readxl::read_xlsx('bg_obs_catch_columns.xlsx')

kable(dat)

```

