---
title: "Groundfish Module Summary"
author: "Ben Galuardi"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    fig_caption: true
  pdf_document: default
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning = FALSE, 
											message = FALSE, cache = FALSE,
											progress = TRUE, verbose = FALSE, comment = F
											, error = FALSE, dev = 'png', dpi = 200)

options(knitr.kable.NA = '')
options(dplyr.summarise.inform = FALSE)
```


### Background 

Discards of groundfish species are used for several purposes throughout the year. Quota Monitoring requires these on a weekly basis. Discard rates for sector trips are shared with Sector managers. End of year ACL accounting also requires discard estimates from all trips. 

This means a full accounting of groundfish discards occurs in several steps. 


### Steps for Groundfish Trips

1. Pull all merged trips from `CAMS_OBS_CATCH`. Here, it is a good idea to pull only a few years and not the whole table, which is ~12 years of data

2. Separate Groundfish trips from non-groundfish trips (Use CAMS `GF` column)

3. Set stratification variables for groundfish trips

```{r, echo = T, eval = F}
# FULL Stratification variables

stratvars = c( 'SPECIES_STOCK'
							# , 'GEARCODE'  # this is the SECGEAR_MAPPED variable
              , 'CAMS_GEAR_GROUP'
              , 'MESHGROUP'
              , 'SECTID'
              , 'EM'  # previously was four columns of EFP information.. 
              , "REDFISH_EXEMPTION"
              , "SNE_SMALLMESH_EXEMPTION"
              , "XLRG_GILLNET_EXEMPTION"
              )
```

	- SPECIES_STOCK is taken from CAMS support table `MAPS.CAMS_STATAREA_STOCK`
	- CAMS_GEAR_GROUP is derived from a support table (`MAPS.CAMS_GEARCODE_STRATA`)
	- MESHGROUP is hardcoded for all trips according to decisions made by the mesh subgroup (see summary when available)
	- SECTID comes from a CAMS matching table (`MAPS.MATCH_MULT_SECTID`)
	- EFP and Exemptions are built in to `CAMS_LANDINGS`, the precurdor for `CAMS_OBS_CATCH`
	
4. Perform *first pass* of `discaRd`
	- there are two sub-passes for year t and year t-1
5. Perform *second pass* of `discaRd` with discard rates rolled up for all Sectors
	- Common Pool is distinguished from the rest of Sectors
	- Simplified stratification is used:
	
```{r, echo = T, eval = F}
# Assumed Stratification variables

stratvars = c('SPECIES_STOCK'
              , 'CAMS_GEAR_GROUP'
              , 'MESHGROUP'
							, 'SECTOR_TYPE'
              ) 
```	

6. The discaRd functions allow for an assumed rate to be calculated. This assumed rate is relative to the stratification used in the functions. Here, we utilize this feature to generate a broad stock rate. the stratification here is simply `SPECIES_STOCK`

7. For each *pass*, a transition rate is calculated between year t and year t-1. This rate determines how much, if any, information is used from previous years. 

8. The two *passes* are joined in a hierarchical manner. Rates and `DISCARD_SOURCE` (in parentheses) are assigned for each trip according to:
 - (I) in season rate; >= 5 trips in Full Stratification
 - (T) Transition in season rate; < 5 trips in Full Stratification, year t, AND >= 5 trips in year t-1
 - (A) Assumed rate. This is the *second pass* rate when there were >=5 trips in season
 - (B) Broad stock rate is used when other criteria are not met. 
 - (O) Observed values used from obserevd trips; discard rate is NOT USED. 
 
9. CV calculations are available for (I), (T), and (A). Obtaining a CV estimate for (B) would require a *third pass* of discaRd functions. (O) rates are not used and final discard values are not estimated. 

10. Discard pounds per trip are calculated according to

```{r echo = T, eval = F}
	mutate(coalesce(DISC_MORT_RATIO, 1)) %>%
	mutate(DISCARD = case_when(!is.na(LINK1) ~ DISC_MORT_RATIO*OBS_DISCARD
														 , is.na(LINK1) ~ DISC_MORT_RATIO*COAL_RATE*LIVE_POUNDS)
				 
			# COAL_RATE is the final discard rate used. It is 'coalesced' from the (I), (A) and (B) rates 	 
```

By assigning `SPECIES_STOCK` as a stratification variable, the computation time is reduced. Each subtrip may only occur in a single statistical area so it should never cross stock boundaries. 

Once the full table (CAMS_OBS_CATCH) is loaded, each species takes ~12 seconds to process on the server.

Output tables are produced for each species. These can easily be recombined. An example table has been shared on `MAPS` and `CAMS_GARFO`

```sql
	MAPS.CAMS_DISCARD_EXAMPLE_GF19
	CAMS_GARFO.CAMS_DISCARD_EXAMPLE_GF19

```

### Diagnostic Plots/Tables

```{r make diagnostic plot, eval = T, echo = F, fig.cap = "Discard Rates by Stock, Species, Discard Source", fig.height = 8, fig.width = 12}

# library(odbc)
# library(RODBC)
library(dplyr, warn.conflicts = FALSE)
library(dbplyr)
library(ggplot2)
# library(config)
library(stringr)
library(discaRd)
library(knitr)
library(kableExtra)
options(scipen = 999)


# condat <- config::get(value = "bgaluardi_cams_garfo", file = "~/config.yml")
# 
# ccon <- dbConnect(odbc::odbc(),
# 									DSN = condat$dsn,
# 									UID = condat$uid,
# 									PWD = condat$pwd)


mapsdat <- config::get(value = "maps", file = "~/config.yml")

bcon <- ROracle::dbConnect(
    drv = ROracle::Oracle(),
    username = mapsdat$uid,
    password = mapsdat$pwd,  
    dbname = "NERO.world"
)

# bcon <- dbConnect(odbc::odbc(), 
# 									DSN = mapsdat$dsn, 
# 									UID = mapsdat$uid, 
# 									PWD = mapsdat$pwd)

db_example = tbl(bcon, sql("

    select *
    from MAPS.CAMS_DISCARD_EXAMPLE_GF2019
     where GF = 1
    AND FED_OR_STATE = 'FED'

")) %>% 
	collect()

# db_example %>% 
# 	filter(DISCARD_SOURCE != 'O') %>% 
# 	group_by(SPECIES_ITIS, DISCARD_SOURCE, STRATA_FULL, STRATA_ASSUMED) %>% 
# 	slice(1) %>% 
# 	ggplot()+
# 	geom_bar(aes(x = SPECIES_STOCK, y = CAMS_DISCARD_RATE, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+
# 	facet_wrap(~COMMON_NAME, scales = 'free')+
# 	theme_light()
```

```{r plot 2, echo = F, fig.cap = "Discard Estimate by Stock, Species, Discard Source" , fig.height = 8, fig.width = 12}
db_example %>% 
	# filter(DISCARD_SOURCE != 'O') %>% 
	group_by(SPECIES_ITIS, COMMON_NAME, DISCARD_SOURCE,SPECIES_STOCK) %>% 
	dplyr::summarise(DSUM = sum(DISCARD, na.rm = T)) %>% 
	# slice(1) %>% 
	ggplot()+
	geom_bar(aes(x = SPECIES_STOCK, y = DSUM, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+
	facet_wrap(~COMMON_NAME, scales = 'free')+
	theme_light()

```


```{r table 1}
db_example %>% 
	group_by(COMMON_NAME
					 , SPECIES_STOCK) %>% 
	dplyr::summarise(nvtr = n_distinct(VTRSERNO)
									 , KALL = sum(SUBTRIP_KALL, na.rm = T)
									 , DISCARD = round(sum(DISCARD, na.rm = T))) %>% 
	DT::datatable(caption = 'Discard Estimate by Stock, Species, Discard Source (using CAMS Stock ID)') %>% 
	DT::formatRound(c('nvtr','KALL','DISCARD'), digits = 0, interval = 3)
	# 
	# knitr::kable(format.args = list(big.mark = ",")) %>% 
	# kable_styling(bootstrap_options = c("striped", "hover", "responsive"))

```

```{r DMIS CAMS DISCARD COMPARISON, echo = F}


match_stockid <- function(bcon, species_itis){
	
	stock_dmis = tbl(bcon, sql("
    select stock_id
    , comname
    , area
    , max(species_itis) as species_itis
    from fso.v_obSpeciesStockArea 
    where stock_id not like 'OTHER'
    and species_itis is not null
    group by stock_id, comname, species_itis, area
")
) %>%
	collect() %>% 
		mutate(SPECIES_ITIS = ifelse(COMNAME == 'OCEAN POUT', "630979", SPECIES_ITIS)) %>% 
		filter(SPECIES_ITIS == species_itis)
	
	 
 t2 = 	tbl(bcon, sql('select * from MAPS.CAMS_STATAREA_STOCK')) %>%
 filter(SPECIES_ITIS == species_itis) %>% 
 	collect() %>%
 	mutate('AREA' = as.character(STAT_AREA))
 
 
 t2 %>% 
 	left_join(., stock_dmis, by = 'AREA') %>% 
 	mutate(CAMS_STOCK_AREA = paste(COMMON_NAME, AREA_NAME, sep = '_')) %>% 
 	dplyr::rename('DMIS_STOCK_ID' = STOCK_ID
 								, 'SPECIES_ITIS' = 'SPECIES_ITIS.x') %>% 
 	dplyr::select(-SPECIES_ITIS.y)
	
	
}


species = tbl(bcon, sql("
select distinct(b.species_itis)
    , COMNAME
    , a.nespp3
from fso.v_obSpeciesStockArea a
left join (select *  from APSD.CAMS_GEARCODE_STRATA) b on a.nespp3 = b.nespp3
where stock_id not like 'OTHER'
and b.species_itis is not null
")
) %>% 
	collect()

discard2019 = tbl(bcon, sql("
select round(sum(POKGMASS_DISCARD)) POKGMASS
,round(sum(CODGMSS_DISCARD)) CODGMSS
,round(sum(CODGBE_DISCARD)) CODGBE
,round(sum(CODGBW_DISCARD)) CODGBW
,round(sum(FLDSNEMA_DISCARD)) FLDSNEMA
,round(sum(FLWGB_DISCARD)) FLWGB
,round(sum(FLWGMSS_DISCARD)) FLWGMSS
,round(sum(PLAGMMA_DISCARD)) PLAGMMA
,round(sum(YELCCGM_DISCARD)) YELCCGM
,round(sum(HADGBW_DISCARD)) HADGBW
,round(sum(WITGMMA_DISCARD)) WITGMMA
-- ,round(sum(FLWGMSS_DISCARD)) FLWGMSS
,round(sum(HALGMMA_DISCARD)) HALGMMA
,round(sum(YELGB_DISCARD)) YELGB
,round(sum(FLGMGBSS_DISCARD)) FLGMGBSS
,round(sum(HKWGMMA_DISCARD)) HKWGMMA
,round(sum(REDGMGBSS_DISCARD)) REDGMGBSS
-- ,round(sum(FLWGB_DISCARD)) FLWGB
,round(sum(HADGM_DISCARD)) HADGM
,round(sum(OPTGMMA_DISCARD)) OPTGMMA
,round(sum(WOLGMMA_DISCARD)) WOLGMMA
,round(sum(FLWSNEMA_DISCARD)) FLWSNEMA
,round(sum(HADGBE_DISCARD)) HADGBE
-- ,round(sum(CODGBW_DISCARD)) CODGBW
,round(sum(YELSNE_DISCARD)) YELSNE

from apsd.dmis_all_years
where fishing_year = 2019
"))  %>% 
  collect() %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate(stock = row.names(.))

names(discard2019)[1] = 'DMIS_DISCARD'

discard2019$STOCK_ID = unlist(lapply(strsplit(discard2019$stock, split = '_'), function(x) x[[1]]))

discard2019 = discard2019 %>% dplyr::select(-stock)


# get CAMS estimates

comp_list = NULL

ii = 1

for(j in species$SPECIES_ITIS){
	
	comp_list[[ii]] <- db_example %>% 
		filter(SPECIES_ITIS == j)
	
	sid = match_stockid(bcon, j) %>% 
		dplyr::select(DMIS_STOCK_ID, AREA, CAMS_STOCK_AREA)
	
	comp_list[[ii]] = comp_list[[ii]] %>% 
		left_join(sid, by = c('AREA')) %>% 
		group_by(COMMON_NAME, CAMS_STOCK_AREA, DMIS_STOCK_ID) %>% 
		dplyr::summarise(CAMS_DISCARD = round(sum(DISCARD, na.rm = T))) %>% 
		filter(!is.na(COMMON_NAME))
	
	ii = ii+1
	
}

allgf_comp = do.call(rbind, comp_list) %>% 
	filter(!is.na(CAMS_STOCK_AREA))

# join to DMIS estimates

dmis_comp = allgf_comp %>% 
	group_by(COMMON_NAME, DMIS_STOCK_ID) %>% 
	dplyr::summarise(CAMS_DISCARD = round(sum(CAMS_DISCARD, na.rm = T))) %>% 
	left_join(., discard2019, by = c('DMIS_STOCK_ID' = 'STOCK_ID')) %>% 
	mutate(DMIS_CAMS_DIFF = DMIS_DISCARD - CAMS_DISCARD
				 , DMIS_CAMS_DIFF_PERC = (DMIS_DISCARD - CAMS_DISCARD)/DMIS_DISCARD)

dmis_comp %>% 
	DT::datatable(caption = 'DISCARD comparison, DMIS/CAMS (Uses DMIS Stock ID for summaries)') %>% 
	DT::formatPercentage('DMIS_CAMS_DIFF_PERC') %>% 
	DT::formatRound(names(dmis_comp)[3:5], digits = 0, interval = 3)



```


```{r DMIS CAMS KALL Comparison}


get_dmis_kall_stock <- function(bcon, species_itis){

 t1 = tbl(bcon, sql("
  select *
  from apsd.dmis_all_years
  where fishing_year = 2019
  and activity_code like 'NMS%'
   "
 )
)
 
 t2 = 	tbl(bcon, sql('select * from MAPS.CAMS_STATAREA_STOCK')) %>%
 filter(SPECIES_ITIS == species_itis)
 
stock_dmis = tbl(bcon, sql("
    select stock_id
    , comname
    , area
    , max(species_itis) as species_itis
    from fso.v_obSpeciesStockArea 
    where stock_id not like 'OTHER'
    and species_itis is not null
    group by stock_id, comname, species_itis, area
")
) %>%
	filter(SPECIES_ITIS == species_itis)



t3 = tbl(bcon, sql("select *
									 from MAPS.CAMS_LANDINGS
									 where activity_code_1 like 'NMS%'
									 AND RECORD_LAND >= '01-MAY-19'
									 AND RECORD_LAND < '01-MAY-20'
									 "
									 )) %>% 
	left_join(., t2, by = c('AREA' = 'STAT_AREA')) %>% 
	filter(!is.na(AREA_NAME)) %>% 
	group_by(AREA_NAME) %>% 
	dplyr::summarise(KALL_CAMS = sum(LIVLB, na.rm = T))

    # collect()
 
 t1 %>% 
 	left_join(., t2, by= c('AREA' = 'STAT_AREA')) %>%
 	left_join(., stock_dmis, by = c('AREA' = 'AREA')) %>% 
 	filter(!is.na(STOCK_ID.y)) %>%   # don't use STOCK_ID.x... species specific stock to attribute catch 
 	group_by(STOCK_ID.y, AREA_NAME) %>%
 	# head() %>% 
 	dplyr::summarise(KALL_DMIS = round(sum(POUNDS, na.rm = T))) %>%
 	left_join(., t3, by = 'AREA_NAME') %>% 
 	collect() %>% 
 	mutate(SPECIES_ITIS = species_itis) %>% 
 	dplyr::rename('STOCK_ID_DMIS' = 'STOCK_ID.y', 'STOCK_ID_CAMS' = 'AREA_NAME')
 
}

# make the table 

species = tbl(bcon, sql("
select distinct(b.species_itis)
    , COMNAME
    , a.nespp3
from fso.v_obSpeciesStockArea a
left join (select *  from APSD.CAMS_GEARCODE_STRATA) b on a.nespp3 = b.nespp3
where stock_id not like 'OTHER'
and b.species_itis is not null
")
) %>% 
	collect()


species$SPECIES_ITIS %>% 
	as.list() %>%
lapply(., function(i) get_dmis_kall_stock(bcon, i)) %>% 
	do.call(rbind, .) %>% 
	mutate(KALL_DIFF = KALL_DMIS - KALL_CAMS
				 , KALL_DIFF_PERC = round((KALL_DMIS - KALL_CAMS)/KALL_DMIS, 3)) %>% 
	DT::datatable(.
								# , extensions = 'Buttons'
							 # , options = list(dom = 'frtipB', buttons = c('csv', 'excel'), autoWidth = TRUE)
							 , filter = "top"
						 	 , class = 'cell-border stripe'
						 	 , rownames = FALSE
							, caption = 'KALL comparison, DMIS/CAMS') %>% 
	DT::formatPercentage('KALL_DIFF_PERC') %>% 
	DT::formatRound(c('KALL_DMIS','KALL_CAMS','KALL_DIFF'), digits = 0, interval = 3)

```

```{r comparison of gear stratifications, eval = F}

load('compare_gear_strata_gftrips19.Rdata')

gearcodes %>% 
		DT::datatable(caption = 'Gears used in OBS/CATCH merge (STG_SECGEAR_MAPPED)')
	# DT::formatRound(names(dmis_comp)[3:5], digits = 0, interval = 3)
	# knitr::kable(format.args = list(big.mark = ",")
	# 						 , caption = "Gears used in OBS/CATCH merge (STG_SECGEAR_MAPPED)") %>% 
	#  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))


CAMS_GEAR_STRATA %>% 
	DT::datatable(caption = 'Gear mapping for discaRd: Ocean Pout example')
	# knitr::kable(format.args = list(big.mark = ",")
	# 						 , caption = "Gear mapping for discaRd: ocean pout example") %>% 
	#  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))



# res_cams_gear %>%
	# replace_na(replace = list(' ')) %>% 
	# 	knitr::kable(format.args = list(big.mark = ",")
	# 							 , caption = "Groundfish discard estimates using CAMS_GEAR_GROUP") %>% 
	#  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))

# res_cams_gearcode %>% 
# 		knitr::kable(format.args = list(big.mark = ",")
# 								 , caption = "Groundfish discard estimates using SECGEAR_MAPPED (OBS/Catch categories)") %>% 
# 	 kable_styling(bootstrap_options = c("striped", "hover", "responsive"))


bind_cols(res_cams_gear %>% dplyr::select(1,2), (res_cams_gear %>% dplyr::select(-1,-2)) - (res_cams_gearcode %>% dplyr::select(-1,-2))) %>% 
		knitr::kable(format.args = list(big.mark = ",")
								 , caption = "Difference in estimates between using CAMS_GEAR_GROUP and SECGEAR_MAPPED gear groupings") %>% 
	 kable_styling(bootstrap_options = c("striped", "hover", "responsive"))


```


### Steps for estimation of groundfish discard on non-Groundfish Trips

1. same as above

2. same as above

3. Set stratification variables

```{r, echo = T, eval = F}
# Full (First Pass) Stratification variables

stratvars = c('SPECIES_STOCK'
              ,'CAMS_GEAR_GROUP'
							, 'MESHGROUP'
						  , 'TRIPCATEGORY'
						  , 'ACCESSAREA')
```

	- SPECIES_STOCK is taken from CAMS support table `MAPS.CAMS_STATAREA_STOCK`
	- CAMS_GEAR_GROUP is derived from a support table (`MAPS.CAMS_GEARCODE_STRATA`)
	- MESHGROUP is hardcoded for all trips according to decisions made by the mesh subgroup (see summary when available)
	- TRIPCATEGORY is specific to scallop trips and indicates Limited Access or General Category Fleets
	- ACCESS AREA is specific to scallop trips and indicates Scallop Access Area

4. Run first pass of `discaRd` for full stratification
	- there are sub-passes for year t and year t-1
	
5. Run a second pass using a coarsened stratification : 'SPECIES_STOCK', 'CAMS_GEAR_GROUP' and 'MESHGROUP'.

	
```{r, echo = T, eval = F}
# Second Pass Stratification variables

stratvars = c('SPECIES_STOCK'
              , 'CAMS_GEAR_GROUP',
              ,'MESHGROUP')

```	


6. Run a third pass using Stock and Gear only
```{r, echo = T, eval = F}
# Third PAss Stratification variables

stratvars = c('SPECIES_STOCK'
              , 'CAMS_GEAR_GROUP',
              ,'MESHGROUP')

```	

7. Trips/Gear strata that are unobserved after these steps are not assigned rates as no observer coverage exists for these gear types (e.g. Menhaden purse seine)

8. Rates and `DISCARD_SOURCE` (in parentheses) are assigned for each trip according to:
 - (I) in season rate; >= 5 trips in Full Stratification
 - (T) Transition in season rate; < 5 trips in Full Stratification, year t, AND >= 5 trips in year t-1
 - (GM) Assignment is from the second pass (e.g. Stock/Gear/Meshgroup). This is the rate when there were <5 trips in season and <5 in previous season at the full stratification level.
 - (G) Assignment is from the second pass (e.g. Stock/Gear). 
 - (N) No coverage in focal season and previous season; discard rate is not possible to estimate.
 - (O) Observed values used from observed rate trips; discard rate is NOT USED. 

9. CV calculations are available for (I), (T), (G) and (GM). 

10. Discard pounds per trip are calculated according to

```{r echo = T, eval = F}
	mutate(DISC_MORT_RATIO = coalesce(DISC_MORT_RATIO, 1)) %>%
	mutate(DISCARD = case_when(!is.na(LINK1) ~ DISC_MORT_RATIO*OBS_DISCARD
														 , is.na(LINK1) ~ DISC_MORT_RATIO*COAL_RATE*LIVE_POUNDS)
				 
	# COAL_RATE is the final discard rate used. It is 'coalesced' from the (I), (GM) and (G) rates 	 
```



```{r plot non groundfish trips, echo = F, fig.cap = "Non-groudfish trips: Discard Estimate by Stock, Species, Discard Source" , fig.height = 8, fig.width = 12}


non_gf = tbl(bcon, sql("

    select SPECIES_ITIS, COMMON_NAME, DISCARD_SOURCE,SPECIES_STOCK
    , sum(DISCARD) as DSUM
    from MAPS.CAMS_DISCARD_EXAMPLE_GF2019
     where GF = 0
     group by SPECIES_ITIS, COMMON_NAME, DISCARD_SOURCE,SPECIES_STOCK

")) %>% 
	collect()

	ggplot(non_gf)+
	geom_bar(aes(x = SPECIES_STOCK, y = DSUM, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+
	facet_wrap(~COMMON_NAME, scales = 'free')+
	theme_light()

```


```{r non-gf trips table}
	
	non_gf = tbl(bcon, sql("

    select COMMON_NAME, SPECIES_STOCK, FED_OR_STATE
    , sum(DISCARD) as DISCARD
    , sum(subtrip_kall) as KALL
    , count(distinct(VTRSERNO)) as N_VTR
    from MAPS.CAMS_DISCARD_EXAMPLE_GF2019
     where GF = 0
     group by COMMON_NAME,SPECIES_STOCK, FED_OR_STATE

")) %>% 
		collect()	

	
	non_gf %>% 
	DT::datatable(caption = 'Non-groundfish trips: Discard Estimate by Stock, Species, Discard Source (using CAMS Stock ID)') %>% 
	DT::formatRound(c('N_VTR','KALL','DISCARD'), digits = 0, interval = 3)

```


<!-- ### To-do  -->

<!-- 1. Check totals between CAMS/DMIS -->
<!-- 	- KALL (**DONE**) -->
<!-- 	- d/k rates () -->

<!-- 2. Complete code for non-groundfish trips in groundfish module. (**DONE**) -->
<!-- 	- combine with previous work (or not) so we have a full set of trips for each groundfish stock (**DONE**) -->

<!-- 3. Apply discard rates to State records (**DONE**) -->
<!--  - this may be done implicitly if applying rates by strata to all CAMS records (e.g. by CAMSID)  -->

<!-- 4. Incorporate EM (including MREM) records is using those values for discard amounts. This should be a matter of substitution on a trip by trip basis. This may yield another `DISCARD_SOURCE` (e.g. (EM)?) -->
<!-- 	- how to deal with adjustment of KALL on MREM trips? -->

<!-- 5. Scallop trip estimates using Scallop Strata. 	(**Done**) -->
<!--  - Scallop Trips: Dan C. does this separately from other trips. Stratifies GF discard rates by - Gear (Trawl/Dredge) - Fleet (LA/LAGC) - does NOT stratify by Access Area/Open; only by stock area - Yellowtail and Windowpane stocks are derived from scallop in season QM procedure (**Done**) -->

<!-- 6. Add CV estimate routine for Assumed strata; Already added this for Broad Stock Rate trips -->

<!-- ### Results/Discussion -->

<!-- - Discard totals for CAMS are now very close to DMIS for groundfish trips.   -->
<!-- - Strata between DMIS and CAMS are not identical. CAMS uses new meshgroupings as well as CAMS_GEAR_GROUP.  -->
<!-- 	- Comparisons between using CAMS_GEAR_GROUP and a more specific gear factorization (SECGEAR_MAPPED) show only slight differences  -->
	
<!-- - One more significant difference between DMIS and CAMS groundfish stratification is that Quota Monitoring (DMIS) uses stock area defined biologically, and parses out the designations (STOCK_ID) in tables above AFTER estimating discards. For example, Eastern Georges Bank (EGB) is CAMS Strata. Although DMIS has a separate total for EGB, this is parsed out from the entire Georges Bank stock. This differential stratification is the most likely explanation for differences between estimates. Stocks that have a singular STOCK_ID (e.g. Pollock), are fairly close in their DMIS/CAMS discard estimates.  -->




