---
title: "Groundfish Module Summary"
author: "Ben Galuardi"
date: "`lubridate::today()`"
output:
  html_document: default
  pdf_document: default
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

# Background 

Discards of groundfish species are used for several purposes throughout the year. Quota Monitoring requires these on a weekly basis. Discard rates for sector trips are shared with Sector managers. End of year ACL accounting also requires discard estimates from all trips. 

This means a full accounting of groundfish discards occurs in several steps. 


### Steps 

1. pull all merged trips from CAMS_OBS_CATCH

2. separate Groundfish trips from non-groundfish trips (Use VMS declaration `like 'NMS%'`)

3. Set stratification variables for groundfish trips

```{r, echo = T, eval = F}
# FULL Stratification variables

stratvars = c('SPECIES_STOCK'
              , 'CAMS_GEAR_GROUP'
              , 'MESHGROUP'
              # , 'CAREA'
              , 'SECTID'
              , "PERMIT_EFP_1"
              , "PERMIT_EFP_2"
              , "PERMIT_EFP_3"
              , "PERMIT_EFP_4"
              , "REDFISH_EXEMPTION"
              , "SNE_SMALLMESH_EXEMPTION"
              , "XLRG_GILLNET_EXEMPTION"
              ) 
```

	- SPECIES_sTOCK is taken from CAMS support table `MAPS.CAMS_STATAREA_STOCK`
	- CAMS_GEAR_GROUP is derived from a support table (`MAPS.CAMS_GEARCODE_STRATA`)
	- MESHGROUP is hardcoded for all trips according to decisions made by the mesh subgroup (see summary when available)
	- SECTID comes from a CAMS matching table (`MAPS.MATCH_MULT_SECTID`)
	- EFP and Exemptions come from the CAMS trip atributes View (`MAPS.STG_TRIP_ATTR`), which complements `MAPS.DLR_VTR`
	
4. Perform *first pass* of `discaRd`
	- there are two sub-passes for year t and year t-1
5. Perform *second pass* of `discaRd` with discard rates rolled up for all Sectors
	- Common Pool is distinguished from the rest of Sectors
	- Simplified stratification is used:
	
```{r, echo = T, eval = F}
# Assumed Stratification variables

stratvars = c('SPECIES_STOCK'
              , 'CAMS_GEAR_GROUP'
              , 'MESHGROUP'
							, 'SECTOR_TYPE'
              ) 
```	

6. The discaRd functions allow for an assumed rate to be calculated. This assumed rate is realtive to the stratification used in the functions. Here, we utilize this feature to generate a broad stock rate. the stratification here is simply `SPECIES_STOCK`

7. For each *pass*, a transition rate is calculated between year t and year t-1. This rate determines how much, if any, information is used from previous years. 

8. The two *passes* are joined in a hierarchical manner. Rates and `DISCARD_SOURCE` (in parentheses) are assigned for each trip according to:
 - (I) in season rate; >= 5 trips in Full Stratification
 - (T) Transition in season rate; < 5 trips in Full Stratification, year t, AND >= 5 trips in year t-1
 - (A) Assumed rate. This is the *second pass* rate when there were >=5 trips in season
 - (B) Broad stock rate is used when other criteria are not met. 
 - (O) Observed values used from obserevd trips; discard rate is NOT USED. 
 
9. CV calculations are available for (I), (T), and (A). Obtaining a CV estimate for (B) would require a *third pass* of discaRd functions. (O) rates are not used and final discard values are not estimated. 

10. Discard pounds per trip are calculated according to

```{r echo = T, eval = F}
	mutate(coalesce(DISC_MORT_RATIO, 1)) %>%
	mutate(DISCARD = case_when(!is.na(LINK1) ~ DISC_MORT_RATIO*OBS_DISCARD
														 , is.na(LINK1) ~ DISC_MORT_RATIO*COAL_RATE*LIVE_POUNDS)
				 
			# COAL_RATE is the final discard rate used. It is 'coalesced' from the (I), (A) and (B) rates 	 
```


By assigning `SPECIES_STOCK` as a stratification variable, the computation time is reduced. Each subtrip may only occur in a single statistical area so it should never cross stock boundaries. 

Once the full table (CAMS_OBS_CATCH) is loaded, each species takes ~12 seconds to process on the server.

Output tables are produced for each species. These can easily be recombined. An example table has been shared on `MAPS` and `CAMS_GARFO`

```sql
	MAPS.CAMS_DISCARD_EXAMPLE_GF19
	CAMS_GARFO.CAMS_DISCARD_EXAMPLE_GF19

```

### Diagnostic Plots/Tables

```{r make diagnostic plot, eval = T, echo = F, fig.cap = "Discard Rates by Stock, Species, Discard Source"}

library(odbc)
library(dplyr, warn.conflicts = FALSE)
library(dbplyr)
library(ggplot2)
# library(config)
library(stringr)
library(discaRd)
options(scipen = 999)

condat <- config::get(value = "bgaluardi_cams_garfo", file = "~/config.yml")

ccon <- dbConnect(odbc::odbc(),
									DSN = condat$dsn,
									UID = condat$uid,
									PWD = condat$pwd)

db_example = tbl(ccon, sql("select * from CAMS_GARFO.CAMS_DISCARD_EXAMPLE_GF19")) %>% 
	collect()

db_example %>% 
	filter(DISCARD_SOURCE != 'O') %>% 
	group_by(SPECIES_ITIS_EVAL, DISCARD_SOURCE, STRATA_FULL, STRATA_ASSUMED) %>% 
	slice(1) %>% 
	ggplot()+
	geom_bar(aes(x = SPECIES_STOCK, y = DISCARD_RATE, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+
	facet_wrap(~COMNAME_EVAL, scales = 'free')+
	theme_light()
```

```{r plot 2, echo = F, fig.cap = "Discard Estimate by Stock, Species, Discard Source"}
db_example %>% 
	# filter(DISCARD_SOURCE != 'O') %>% 
	group_by(SPECIES_ITIS_EVAL, COMNAME_EVAL, DISCARD_SOURCE,SPECIES_STOCK) %>% 
	dplyr::summarise(DSUM = sum(DISCARD, na.rm = T)) %>% 
	# slice(1) %>% 
	ggplot()+
	geom_bar(aes(x = SPECIES_STOCK, y = DSUM, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+
	facet_wrap(~COMNAME_EVAL, scales = 'free')+
	theme_light()

```

```{r table 1}
db_example %>% 
	group_by(COMNAME_EVAL
					 , SPECIES_STOCK) %>% 
	dplyr::summarise(nvtr = n_distinct(VTRSERNO)
									 , KALL = sum(SUBTRIP_KALL, na.rm = T)
									 , DISCARD = round(sum(DISCARD, na.rm = T))) %>% 
	knitr::kable(format.args = list(big.mark = ","))

```

```{r discard rates by strata example}

db_example %>% 
	group_by(COMNAME_EVAL
					 , DISCARD_SOURCE
					 , STRATA_FULL
					 , STRATA_ASSUMED) %>% 
	dplyr::summarise(DISCARD_RATE = max(DISCARD_RATE)
									 , KALL = sum(SUBTRIP_KALL, na.rm = T)) %>% 
	knitr::kable(format.args = list(big.mark = ","))


```



### Issues

### To-do 

1. complete code for non-groundifsh trips in groundfish module. 
	- combine with previous work (or not) so we have a full set of trips for each groundfish stock
	- Scallop Trips: Dan C. does this separately from other trips. Stratifies GF discard rates by - Gear (Trawl/Dredge) - Fleet (LA/LAGC) - does NOT stratify by Access Area/Open; only by stock area - Yellowtail and Windowpane stocks are derived from scallop in season QM procedure
2. Apply discard rates to State trips	
3. Incorporate EM (including MREM) records is using those values for discard amounts. This shoudl be a matter of substitution on a trip by trip basis. This may yield another `DISCARD_SOURCE` (e.g. (EM)?)


#### OCEAN POUT example

Difference in stratification

```sql
-- number of gear groups for Ocean Pout
select distinct(CAMS_GEAR_GROUP)
from APSD.CAMS_GEARCODE_STRATA
where NESPP3 = 250
```

| CAMS\_GEAR\_GROUP | Gear Type               |
|:------------------|:------------------------|
| 0                 | other                   |
| 100               | Gillnet                 |
| 132               | scallop dredge          |
| 50                | Trawls                  |
| 54                | Ruhle Trawl             |
| 57                | Haddock Separator Trawl |

Dan Caless's summary has nine gear groupings:

Gillnet Longline and other line gear Otter trawl Pot, lobster Pair trawl Purse seine Pot and traps Twin trawl Other



## Feb 1, 2022

-   Multiple LINK1 on a single VTR are sneaking through our fix using only observed hauls. The species found on these trips, for the entire 2018-2020 dataset in groundfish tips are : \] "660" "667" "679" "680" "681" "682" "683" "685" "687" "689" NA which correspond to debris unknown fish groups, random invertebrates etc. see following:

```sql 
select *
from obdbs.obspec@NOVA
where substr(NESPP4,1,3) in (660
  , 667
  , 679
  , 680
  , 681
  , 682
  , 683
  , 685
  , 687
  , 689)
```

There are no `SPECIES ITIS` codes for these `NESPP3` codes and are most likely not ever estimated.

Solution : this has been fixed at the SPECIES_ITIS level


## Feb 3, 2022

- TRIPEXT should be only C and X.. DMIS discards are calculated from a table provided by the observer program; and they only use these trip designations.
- Examined Dan Caless's discard script `Discard_Calculations_21_Mults.sql` this is 1500 lines of nested tables and exceptions..
- should re-evaluate whether a new groundfish module is necessary or whether porting the SQL to CAMS is a better option.
- the final discard rate table is not a bad template for an output table.

## Feb 4, 2022

- Removing TRIP_EXT != C or X removes 30,000 of ~80,000 link1 for the entire OBS matching table.

- This should be done in the `MERGE_CAMS_OBS.sql` stage

#### Comparison of results for Haddock 2019

- discard rates for EGB Haddock were ~ 4x lower using CAMS approach than DMIS approach.

- There were ~12 strata in DMIS that did not show up in CAMS as being in season (>=5 trips in the strata). These all fell into the `Trawl LM` category.

- After controlling other aspects, this is likely due to the `CAMS_GEAR_GROUP` and `MESHGROUP` used in CAMS.

- There is a possibility that CAMS trips are not matching to OBS trips on multi-VTR trips due to one element (e.g. AREA) not matching. This is less likely however than the previous point.  

```sql
select distinct(discard_rate)
--,disc_rate_type
, sector_id
, secgearfish
, mesh_cat
from fso.t_observer_discard_rate_priv
where fishing_year = 2019
and STOCK_ID = 'HADGBE'
and disc_rate_type = 'I'

```


| DISCARD_RATE | SECTOR_ID | SECGEARFISH | MESH_CAT |
|--------------|-----------|-------------|----------|
| 0.019379591  | 16        | OTB         | LM       |
| 0.019379591  | 16        | OTF         | LM       |
| 0.036318597  | 22        | OTF         | LM       |
| 0.019379591  | 16        | OTC         | LM       |
| 0.076209491  | 22        | OTB         | LM       |
| 0.076209491  | 22        | OTF         | LM       |
| 0.019379591  | 16        | OTH         | LM       |
| 0.036318597  | 22        | OTC         | LM       |
| 0.076209491  | 22        | OTC         | LM       |
| 0.036318597  | 22        | OTH         | LM       |
| 0.036318597  | 22        | OTB         | LM       |
| 0.076209491  | 22        | OTH         | LM       |

Table of DMIS discards for HADGBE 2019 strata not in CAMS

```SQL
select distinct(discard_rate)
,disc_rate_type
from fso.t_observer_discard_rate_priv
where fishing_year = 2019
and STOCK_ID = 'HADGBE'
and disc_rate_type <> 'I'

```

| DISCARD_RATE | DISC_RATE_TYPE |
|--------------|----------------|
| 0.046521148  | A              |
| 0            | A              |
| 0.057787552  | A              |
| 0.034521631  | A              |
| 0.035437883  | T              |
| 0.021208026  | T              |
| 0.032564803  | T              |
| 0.044642809  | T              |
| 0.040299341  | T              |
| 0.033404436  | T              |
| 0.024165142  | T              |

unique rates, either Assumed (A) or Transition (T) from DMIS. Different combinations of gear and mesh may share rates; this is all unique values.

|SECTOR_TYPE |DISCARD_SOURCE |CAMS_GEAR_GROUP |MESHGROUP |     drate|
|:-----------|:--------------|:---------------|:---------|---------:|
|COMMON_POOL |A              |50              |LM        | 0.0000000|
|COMMON_POOL |B              |0               |na        | 0.0003783|
|COMMON_POOL |B              |10              |na        | 0.0005245|
|COMMON_POOL |B              |100             |LM        | 0.0027069|
|COMMON_POOL |B              |100             |XL        | 0.0000105|
|COMMON_POOL |B              |50              |LM        | 0.0006650|
|COMMON_POOL |B              |50              |SM        | 0.0000000|
|SECTOR      |A              |100             |LM        | 0.0006698|
|SECTOR      |A              |100             |XL        | 0.0000776|
|SECTOR      |A              |50              |LM        | 0.0066887|
|SECTOR      |A              |57              |LM        | 0.0301717|
|SECTOR      |AT             |0               |na        | 0.0005245|
|SECTOR      |AT             |10              |na        | 0.0203849|
|SECTOR      |B              |0               |na        | 0.0005402|
|SECTOR      |B              |10              |na        | 0.0004913|
|SECTOR      |B              |100             |LM        | 0.0032755|
|SECTOR      |B              |100             |XL        | 0.0006957|
|SECTOR      |B              |132             |na        | 0.0005245|
|SECTOR      |B              |57              |LM        | 0.0301705|

CAMS Assumed (A, AT) and Broad Stock (B) discard rates for EGB Haddock 2019. Generally, these are much lower than the Assumed rates in DMIS.

```r
# Get totals by Stock for Haddock 2019

joined_table = joined_table %>%
	mutate(DISCARD = case_when(!is.na(LINK1) ~ DISC_MORT_RATIO*OBS_DISCARD
														 , is.na(LINK1) ~ DISC_MORT_RATIO*COAL_RATE*LIVE_POUNDS)
				 )

joined_table %>%
	group_by(SPECIES_STOCK, DISCARD_SOURCE) %>%
	dplyr::summarise(DISCARD_EST = sum(DISCARD)) %>%
	pivot_wider(names_from = 'SPECIES_STOCK', values_from = 'DISCARD_EST') %>%
	dplyr::select(-1) %>%
	colSums(na.rm = T)

```

|         SPECIES_STOCK     |        Discard|
|:-------------|---------:|
|EGB           |  25,841|
|GOM           | 150,784|
|MA            |      0.00|
|WGB and South | 309,689|

CAMS discard estimate for Haddock 2019  


| STOCK  | SPECIES | DISCARD ESTIMATE |
|--------|---------|------------------|
| HADGBE | HADDOCK |         107,176  |
| HADGBW | HADDOCK |         384,367  |
| HADGM  | HADDOCK |         202,985  |

DMIS estimate for Haddock 2019 

- EGB shows the highest % difference (~75%) while the others show ~25% difference.
-