---
title: "November Year Module Summary"
#author: "Ben Galuardi modified by K.Winiarski"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    fig_caption: true
  pdf_document: default
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning = FALSE, 
											message = FALSE, cache = FALSE,
											progress = TRUE, verbose = FALSE, comment = F
											, error = FALSE, dev = 'png', dpi = 200)

options(knitr.kable.NA = '')
```


### Background 

Exploration of discard estimates from November year module 

### Steps 

1. pull all merged trips from CAMS_OBS_CATCH

2. Set stratification variables for all trips
```{r, echo = T, eval = F}
# FULL Stratification variables

stratvars = c('SPECIES_STOCK'
              ,'CAMS_GEAR_GROUP'
							, 'MESHGROUP'
						  , 'TRIPCATEGORY'
						  , 'ACCESSAREA')
```

	- SPECIES_STOCK is taken from CAMS support table `MAPS.CAMS_STATAREA_STOCK`
	- CAMS_GEAR_GROUP is derived from a support table (`MAPS.CAMS_GEARCODE_STRATA`)
	- MESHGROUP is hardcoded for all trips according to decisions made by the mesh subgroup (see summary when available)
	- SECTID comes from a CAMS matching table (`MAPS.MATCH_MULT_SECTID`)

3. Run `discaRd`
	- there are sub-passes for year t and year t-1
	
```{r, echo = T, eval = F}
# Assumed Stratification variables

stratvars = c('SPECIES_STOCK'
              , 'CAMS_GEAR_GROUP')

```	

4. The discaRd functions allow for an assumed rate to be calculated. This assumed rate is realtive to the stratification used in the functions. Here, the stratification is coarsened to 'SPECIES_STOCK' and 'CAMS_GEAR_GROUP'

7. A transition rate is calculated between year t and year t-1. This rate determines how much, if any, information is used from previous years. 

8. Rates and `DISCARD_SOURCE` (in parentheses) are assigned for each trip according to:
 - (I) in season rate; >= 5 trips in Full Stratification
 - (T) Transition in season rate; < 5 trips in Full Stratification, year t, AND >= 5 trips in year t-1
 - (A) Assumed rate. This is the rate when there were <5 trips in season and <5 in previous season.
 - (O) Observed values used from obserevd trips; discard rate is NOT USED. 
9. CV calculations are available for (I) and (T). Obtaining a CV estimate for (A) would require a *second pass* of discaRd functions.

10. Discard pounds per trip are calculated according to

```{r echo = T, eval = F}
	mutate(DISC_MORT_RATIO = coalesce(DISC_MORT_RATIO, 1)) %>%
	mutate(DISCARD = case_when(!is.na(LINK1) ~ DISC_MORT_RATIO*OBS_DISCARD
														 , is.na(LINK1) ~ DISC_MORT_RATIO*COAL_RATE*LIVE_POUNDS)
				 
			# COAL_RATE is the final discard rate used. It is 'coalesced' from the (I), (A) and (B) rates 	 
```

By assigning `SPECIES_STOCK` as a stratification variable, the computation time is reduced. Each subtrip may only occur in a single statistical area so it should never cross stock boundaries. 

Once the full table (CAMS_OBS_CATCH) is loaded, each species takes ~X seconds to process on the server.

Output tables are produced for each species. These can easily be recombined. Preliminary species tables have been shared on `MAPS` and `CAMS_GARFO`.  

```sql
	MAPS.CAMS_DISCARD_EXAMPLE_CY_GOLDENTILEFISH_19
	CAMS_GARFO.CAMS_DISCARD_EXAMPLE_CY_GOLDENTILEFISH_19
	
	MAPS.CAMS_DISCARD_EXAMPLE_CY_BLUELINETILEFISH_19
	CAMS_GARFO.CAMS_DISCARD_EXAMPLE_CY_BLUELINETILEFISH_19

```

### Diagnostic Plots/Tables

```{r make diagnostic plot, eval = T, echo = F, fig.cap = "2019 Federal Discard Totals by Source", fig.height = 8, fig.width = 12}

# library(odbc)
# library(dplyr, warn.conflicts = FALSE)
# library(dbplyr)
library(ggplot2)
# # library(config)
# library(stringr)
# library(discaRd)
# library(knitr)
# library(kableExtra)
devtools::load_all()
options(scipen = 999)

discard <- read.csv("~/discaRd/CAMS/MODULES/NOVEMBER/November_comparisons.csv")

'%ni%' <- Negate("%in%")

discard <- discard %>% filter(Source %in% c("Federal"))# %>% filter(Species %ni% c("Summer flounder"))


discard %>% 
	ggplot()+
	geom_bar(aes(x = Method, y = Discards), stat = 'identity', position = 'dodge')+
	facet_wrap(~Species, scales = 'free')+
	theme_light()

```

  


```{r make diagnostic plot discards, eval = T, echo = F, fig.cap = "2019 CAMS State Discard Totals", fig.height = 8, fig.width = 12}

discard <- read.csv("~/discaRd/CAMS/MODULES/NOVEMBER/November_comparisons.csv")

'%ni%' <- Negate("%in%")

discard %>% filter(Source %in% c("State")) %>% dplyr::select(Species, Discards) %>%
	DT::datatable(caption = 'Total discards by species for state trips')

```




```{r make diagnostic plot kall, eval = T, echo = F, fig.cap = "2019 KALL Totals by Method and Source", fig.height = 8, fig.width = 12}

Kall <- read.csv("~/discaRd/CAMS/MODULES/NOVEMBER/November_kall.csv")

Kall %>% 
	ggplot()+
	geom_bar(aes(x = Method, y = kall, fill = Source), stat = 'identity', position = 'dodge')+
	#facet_wrap(~Species, scales = 'free')+
	theme_light()

```

Kall is higher for CAMS but due to differences in what species are included in landings (e.g., no clams in DMIS). 


```{r make diagnostic mackerel plot, eval = T, echo = F, fig.cap = "Histogram of Discard Rates by Species, Strata and Discard Source", fig.height = 8, fig.width = 12}

 dw_maps <- config::get(config = "maps", file = "~/config_group.yml")
 
# Connect to database - move this to config file in the future - quick addition for server
  connectString <- paste(
    "(DESCRIPTION=",
    "(ADDRESS=(PROTOCOL=tcp)(HOST=", dw_maps$host, ")(PORT=", dw_maps$port, "))",
    "(CONNECT_DATA=(SERVICE_NAME=",dw_maps$svc, ")))",
    sep = ""
  )

# Connect to oracle each loop in case of timeouts
    bcon <- ROracle::dbConnect(
      drv = ROracle::Oracle(),
      username = dw_maps$uid,
      password = dw_maps$pwd,
      dbname = connectString
    )

db_goldentilefish = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_EXAMPLE_CY_GOLDENTILEFISH_19")) %>% 
	collect

db_goldentilefish$SPECIES <- "GOLDENTILEFISH"

db_bluelinetilefish = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_EXAMPLE_CY_BLUELINETILEFISH_19")) %>%
	collect

db_bluelinetilefish$SPECIES <- "BLUELINETILEFISH"


db_final <- rbind(db_goldentilefish, db_bluelinetilefish)

strat_goldentilefish <- db_goldentilefish %>% 	
  filter(DISCARD_SOURCE != 'O') %>%
  group_by(SPECIES, STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(
                              SPECIES,
                              STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,TRIP_TYPE)
             # ,CV) 
             #dplyr::arrange(desc(DISCARD))

strat_bluelinetilefish <- db_bluelinetilefish %>%
  filter(DISCARD_SOURCE != 'O') %>%
  group_by(SPECIES, STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>%
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(
                              SPECIES,
                              STRATA
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,TRIP_TYPE)

strat_final <- rbind(strat_goldentilefish, strat_bluelinetilefish)
  
  
strat_final %>% 
	filter(DISCARD_SOURCE != 'O' & CAMS_DISCARD_RATE) %>% 
	dplyr::group_by(SPECIES, STRATA, DISCARD_SOURCE, TRIP_TYPE) %>% 
	#slice(1) %>% 
	ggplot()+
	geom_histogram(aes(x = CAMS_DISCARD_RATE, fill = DISCARD_SOURCE), bins=60, position = "dodge") +
	facet_wrap(~SPECIES)+
	theme_light()
```



```{r plot 1, echo = F, fig.cap = "Total Discard Estimates by Species, Stock and Discard Source" , fig.height = 8, fig.width = 12}
db_final %>% 
	# filter(DISCARD_SOURCE != 'O') %>% 
	group_by(SPECIES, SPECIES_ITIS, DISCARD_SOURCE,SPECIES_STOCK) %>% 
	dplyr::summarise(DSUM = sum(DISCARD, na.rm = T)) %>% 
	# slice(1) %>% 
	ggplot()+
	geom_bar(aes(x = SPECIES_STOCK, y = DSUM, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+
 facet_wrap(~SPECIES, scales = 'free')+
	theme_light()
```

Not suprisingly, total discards at the end of the year come from strata with an in season rate. 


```{r goldentilefish discard table }

db_goldentilefish %>% 	
  filter(DISCARD_SOURCE != 'O' & TRIP_TYPE=="FED") %>%
  group_by(STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
  mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>% top_n(10) %>%
		DT::datatable(caption = 'Discard rates by Strata for Golden Tilefish')
```



```{r bluelinetilefish discard table }

  db_bluelinetilefish %>%
   filter(DISCARD_SOURCE != 'O' & TRIP_TYPE=="FED") %>%
   group_by(STRATA_FULL, STRATA_ASSUMED, DISCARD_SOURCE, TRIP_TYPE) %>%
   dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                        , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                        , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                       # , CV = mean(CV, na.rm=TRUE)
                        , DISCARD = round(sum(DISCARD))) %>%
                       # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>%
   mutate(STRATA = case_when(DISCARD_SOURCE == "I" ~ STRATA_FULL, DISCARD_SOURCE == "A" ~ STRATA_ASSUMED)) %>% ungroup() %>% dplyr::select(STRATA
              # ,N_OBS_TRIPS_F
               ,CAMS_DISCARD_RATE
               ,SUBTRIP_KALL
               ,DISC_MORT_RATIO
               ,DISCARD) %>%
              # ,CV)
              dplyr::arrange(desc(DISCARD)) %>% top_n(10) %>%
 		DT::datatable(caption = 'Discard rates by Strata for Blueline Tilefish based on federal trips')
```


### Appendix

For more information regarding stratfication variables and discard rates by species see

MAPS.CAMS_STATAREA_STOCK
MAPS.CAMS_DISCARD_MORTALITY_STOCK
MAPS.CAMS_GEARCODE_STRATA
MAPS.CAMS_MASTER_GEAR
MAPS.CAMS_OBS_CATCH

I may need to grant selection to APSD if you do not have access to MAPS.