---
title: "May Year Module Summary"
#author: "Ben Galuardi modified by K.Winiarski"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    fig_caption: true
  pdf_document: default
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning = FALSE, 
											message = FALSE, cache = FALSE,
											progress = TRUE, verbose = FALSE, comment = F
											, error = FALSE, dev = 'png', dpi = 200)

options(knitr.kable.NA = '')
```


### Background 

Exploration of discard estimates from may year module 

### Steps 

1. pull all merged trips from CAMS_OBS_CATCH

2. Set stratification variables for all trips

```{r, echo = T, eval = F}
# FULL Stratification variables

stratvars = c('SPECIES_STOCK'
              ,'CAMS_GEAR_GROUP'
							, 'MESHGROUP'
						  , 'TRIPCATEGORY'
						  , 'ACCESSAREA')
```

	- SPECIES_STOCK is taken from CAMS support table `MAPS.CAMS_STATAREA_STOCK`
	- CAMS_GEAR_GROUP is derived from a support table (`MAPS.CAMS_GEARCODE_STRATA`)
	- MESHGROUP is hardcoded for all trips according to decisions made by the mesh subgroup (see summary when available)
	- SECTID comes from a CAMS matching table (`MAPS.MATCH_MULT_SECTID`)

3. Run `discaRd`
	- there are sub-passes for year t and year t-1
	
```{r, echo = T, eval = F}
# Assumed Stratification variables

stratvars = c('SPECIES_STOCK'
              , 'CAMS_GEAR_GROUP',
              ,'MESHGROUP')

```	

4. The discaRd functions allow for an assumed rate to be calculated. This assumed rate is relative to the stratification used in the functions. Here, the stratification is coarsened to 'SPECIES_STOCK', 'CAMS_GEAR_GROUP'and 'MESHGROUP'.

7. A transition rate is calculated between year t and year t-1. This rate determines how much, if any, information is used from previous years. 

8.  A broad stock/gear stratification is applied to subtrips that do not meet the requirements for the full or assumed stratification.  This broad stock/gear stratification utilizes observed coverage from both the focal year and the previous year as it includes gear types with low observer coverage.

8. Rates and `DISCARD_SOURCE` (in parentheses) are assigned for each trip according to:
 - (I) in season rate; >= 5 trips in Full Stratification
 - (T) Transition in season rate; < 5 trips in Full Stratification, year t, AND >= 5 trips in year t-1
 - (A) Assumed rate. This is the rate when there were <5 trips in season and <5 in previous season.
 - (G) Broad stock/gear rate <5 trips in season and <5 trips in previous season 
 - (N) No coverage in focal season and previous season; discard rate is not possible to estimate.
 - (O) Observed values used from observed rate trips; discard rate is NOT USED. 

9. CV calculations are available for (I), (T), (A) and (G). 

10. Discard pounds per trip are calculated according to

```{r echo = T, eval = F}
	mutate(DISC_MORT_RATIO = coalesce(DISC_MORT_RATIO, 1)) %>%
	mutate(DISCARD = case_when(!is.na(LINK1) ~ DISC_MORT_RATIO*OBS_DISCARD
														 , is.na(LINK1) ~ DISC_MORT_RATIO*COAL_RATE*LIVE_POUNDS)
				 
			# COAL_RATE is the final discard rate used. It is 'coalesced' from the (I), (A) and (G) rates 	 
```

By assigning `SPECIES_STOCK` as a stratification variable, the computation time is reduced. Each subtrip may only occur in a single statistical area so it should never cross stock boundaries. 

Once the full table (CAMS_OBS_CATCH) is loaded, each species takes ~X seconds to process on the server.

Output tables are produced for each species. These can easily be recombined. Preliminary species tables have been shared on `MAPS` and `CAMS_GARFO`.  

```sql
	MAPS.CAMS_DISCARD_DOGFISHSPINY_2019
	CAMS_GARFO.CAMS_DISCARD_DOGFISHSPINY_2019
	
	MAPS.CAMS_DISCARD_MONKFISH_2019
	CAMS_GARFO.CAMS_DISCARD_MONKFISH_2019
	
	MAPS.CAMS_DISCARD_OFFSHOREHAKE_2019
	CAMS_GARFO.CAMS_DISCARD_OFFSHOREHAKE_2019
	
	MAPS.CAMS_DISCARD_REDHAKE_2019
	CAMS_GARFO.CAMS_DISCARD_REDHAKE_2019
	
	MAPS.CAMS_DISCARD_SILVERHAKE_2019
	CAMS_GARFO.CAMS_DISCARD_SILVERHAKE_2019


```

### Diagnostic Plots/Tables

```{r make diagnostic plot, eval = T, echo = F, fig.cap = "2019 Federal Discard Totals by Source", fig.height = 8, fig.width = 12}

# library(odbc)
# library(dplyr, warn.conflicts = FALSE)
# # library(dbplyr)
library(ggplot2)
# # library(config)
# library(stringr)
# library(discaRd)
# library(knitr)
# library(kableExtra)
devtools::load_all()
options(scipen = 999)

discard <- read.csv("~/discaRd/CAMS/MODULES/MAY/May_comparisons.csv")

'%ni%' <- Negate("%in%")

discard %>% filter(Source == 'Federal') %>%  pivot_wider(id_cols = !Source, names_from = Method,values_from =  c(Discards, kall)) %>% mutate(diff = Discards_CAMS-Discards_QM, perc_diff = 100 * (Discards_QM - Discards_CAMS)/(Discards_CAMS)) %>% dplyr::select(Species, 
Discards_CAMS, Discards_QM, diff, perc_diff) %>%dplyr::mutate(across(where(is.numeric), round, 2))%>% 
		DT::datatable(caption = 'Comparison of total federal discards in FY 2019 by species and source')

```

 

```{r make diagnostic plot discards, eval = T, echo = F, fig.cap = "2019 CAMS State Discard Totals", fig.height = 8, fig.width = 12}

discard <- read.csv("~/discaRd/CAMS/MODULES/MAY/May_comparisons.csv")

'%ni%' <- Negate("%in%")

discard %>% filter(Source %in% c("State")) %>% dplyr::select(Species, Discards) %>%
	DT::datatable(caption = 'Total discards by species for state trips')

```


```{r make diagnostic mackerel plot, eval = T, echo = F, fig.cap = "Histogram of Discard Rates by Species, Strata and Discard Source for Federal trips", fig.height = 8, fig.width = 12}

 dw_maps <- config::get(config = "maps", file = "~/config_group.yml")
 
# Connect to database - move this to config file in the future - quick addition for server
  connectString <- paste(
    "(DESCRIPTION=",
    "(ADDRESS=(PROTOCOL=tcp)(HOST=", dw_maps$host, ")(PORT=", dw_maps$port, "))",
    "(CONNECT_DATA=(SERVICE_NAME=",dw_maps$svc, ")))",
    sep = ""
  )

# Connect to oracle each loop in case of timeouts
    bcon <- ROracle::dbConnect(
      drv = ROracle::Oracle(),
      username = dw_maps$uid,
      password = dw_maps$pwd,
      dbname = connectString
    )

db_dogfish = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_DOGFISHSPINY_2019")) %>% 
	collect

db_dogfish$SPECIES <- "DOGFISH"

db_offshore = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_OFFSHOREHAKE_2019")) %>% 
	collect

db_offshore$SPECIES <- "OFFSHORE HAKE"

db_red = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_REDHAKE_2019")) %>% 
	collect

db_red$SPECIES <- "RED HAKE"

db_silver = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_SILVERHAKE_2019")) %>% 
	collect

db_silver$SPECIES <- "SILVER HAKE"

db_monkfish = tbl(bcon, sql("select * from MAPS.CAMS_DISCARD_MONKFISH_2019")) %>%
	collect

db_monkfish$SPECIES <- "MONKFISH"


db_final <- rbind(db_dogfish, db_monkfish, db_offshore, db_red, db_silver)

strat_dogfish <- db_dogfish %>% 	
   dplyr::group_by(SPECIES, STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                   dplyr::select(
                              SPECIES
                              ,STRATA_FULL
                             # STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,FED_OR_STATE)

strat_offshore <- db_offshore %>% 	
    dplyr::group_by(SPECIES, STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                   dplyr::select(
                              SPECIES
                              ,STRATA_FULL
                             # STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,FED_OR_STATE)


strat_red <- db_red %>% 	
  dplyr::group_by(SPECIES, STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                   dplyr::select(
                              SPECIES
                              ,STRATA_FULL
                             # STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,FED_OR_STATE)



strat_silver <- db_silver %>% 	
 dplyr::group_by(SPECIES, STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                   dplyr::select(
                              SPECIES
                              ,STRATA_FULL
                             # STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,FED_OR_STATE)


strat_monkfish <- db_monkfish %>%
  dplyr::group_by(SPECIES, STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                   dplyr::select(
                              SPECIES
                              ,STRATA_FULL
                             # STRATA 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD
              ,DISCARD_SOURCE
              ,FED_OR_STATE)


strat_final <- rbind(strat_dogfish, strat_monkfish, strat_offshore,
                     strat_red, strat_silver)
  
  
  strat_final %>% 
	filter(FED_OR_STATE=='FED') %>% 
#	dplyr::group_by(SPECIES, STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>% 
	#slice(1) %>% 
	ggplot()+
	geom_histogram(aes(x = CAMS_DISCARD_RATE, fill = DISCARD_SOURCE), bins=60, position = "dodge") +
	facet_wrap(~SPECIES)+
	theme_light()

```

```{r make diagnostic state plot, eval = T, echo = F, fig.cap = "Histogram of Discard Rates by Species, Strata and Discard Source for State trips", fig.height = 8, fig.width = 12}

strat_final %>% 
	filter(DISCARD_SOURCE != 'O' & FED_OR_STATE=='STATE') %>% 
	#dplyr::group_by(SPECIES, STRATA, DISCARD_SOURCE, FED_OR_STATE) %>% 
	#slice(1) %>% 
	ggplot()+
	geom_histogram(aes(x = CAMS_DISCARD_RATE, fill = DISCARD_SOURCE), bins=60, position = "dodge") +
	facet_wrap(~SPECIES)+
	theme_light()
```


```{r plot 1, echo = F, fig.cap = "Total Discard Estimates by Species, Stock and Discard Source for Federal trips" , fig.height = 8, fig.width = 12}
db_final %>% 
	# filter(DISCARD_SOURCE != 'O')&
  filter(FED_OR_STATE=='FED') %>% 
	group_by(SPECIES, SPECIES_ITIS, DISCARD_SOURCE,SPECIES_STOCK) %>% 
	dplyr::summarise(DSUM = sum(CAMS_DISCARD, na.rm = T)) %>% 
	# slice(1) %>% 
	ggplot()+
	geom_bar(aes(x = SPECIES_STOCK, y = DSUM, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+
 facet_wrap(~SPECIES, scales = 'free')+
	theme_light()
```



```{r plot 2, echo = F, fig.cap = "Total Discard Estimates by Species, Stock and Discard Source for State trips" , fig.height = 8, fig.width = 12}
db_final %>% 
	# filter(DISCARD_SOURCE != 'O')&
  filter(FED_OR_STATE=='STATE') %>% 
	group_by(SPECIES, SPECIES_ITIS, DISCARD_SOURCE,SPECIES_STOCK) %>% 
	dplyr::summarise(DSUM = sum(CAMS_DISCARD, na.rm = T)) %>% 
	# slice(1) %>% 
	ggplot()+
	geom_bar(aes(x = SPECIES_STOCK, y = DSUM, fill = DISCARD_SOURCE), stat = 'identity', position = 'dodge')+
 facet_wrap(~SPECIES, scales = 'free')+
	theme_light()
```

```{r dogfish federal discard table }

db_dogfish %>% 	
  filter(DISCARD_SOURCE != 'O' & FED_OR_STATE=="FED") %>%
  group_by(STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
   ungroup() %>% dplyr::select(STRATA_FULL,
                                                                                                                                          DISCARD_SOURCE 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>%# top_n(10) %>%
		DT::datatable(caption = 'Discard rates for federal trips by strata for Dogfish')

```


```{r dogfish state discard table }

db_dogfish %>% 	
  filter(DISCARD_SOURCE != 'O' & FED_OR_STATE=="STATE") %>%
  group_by(STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
   ungroup() %>% dplyr::select(STRATA_FULL,
                                                                                                                                          DISCARD_SOURCE 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>%# top_n(10) %>%
		DT::datatable(caption = 'Discard rates for state trips by strata for Dogfish')

```


```{r monkfish fed discard table }

db_monkfish %>% 	
   filter(DISCARD_SOURCE != 'O' & FED_OR_STATE=="FED") %>%
  group_by(STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
   ungroup() %>% dplyr::select(STRATA_FULL,
                                                                                                                                          DISCARD_SOURCE 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>%# top_n(10) %>%
		DT::datatable(caption = 'Discard rates by Strata for Monkfish')
```

```{r monkfish state discard table }

db_monkfish %>% 	
   filter(DISCARD_SOURCE != 'O' & FED_OR_STATE=="STATE") %>%
  group_by(STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
   ungroup() %>% dplyr::select(STRATA_FULL,
                                                                                                                                          DISCARD_SOURCE 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>%# top_n(10) %>%
		DT::datatable(caption = 'Discard rates by Strata for Monkfish')
```

```{r offshore hake federal discard table }

db_offshore %>% 	
  filter(DISCARD_SOURCE != 'O' & FED_OR_STATE=="FED") %>%
  group_by(STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
   ungroup() %>% dplyr::select(STRATA_FULL,
                                                                                                                                          DISCARD_SOURCE 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>%# top_n(10) %>%
		DT::datatable(caption = 'Discard rates for federal trips by strata for Offshore Hake')
```

```{r offshore hake state discard table }

db_offshore %>% 	
  filter(DISCARD_SOURCE != 'O' & FED_OR_STATE=="STATE") %>%
  group_by(STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
   ungroup() %>% dplyr::select(STRATA_FULL,
                                                                                                                                          DISCARD_SOURCE 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>%# top_n(10) %>%
		DT::datatable(caption = 'Discard rates for state trips by strata for Offshore Hake')
```

```{r redhake federal discard table }

db_red %>% 	
  filter(DISCARD_SOURCE != 'O' & FED_OR_STATE=="FED") %>%
  group_by(STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
   ungroup() %>% dplyr::select(STRATA_FULL,
                                                                                                                                          DISCARD_SOURCE 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>%# top_n(10) %>%
		DT::datatable(caption = 'Discard rates for federal trips by strata for Red Hake')
```

```{r redhake state discard table }

db_red %>% 	
  filter(DISCARD_SOURCE != 'O' & FED_OR_STATE=="STATE") %>%
  group_by(STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
   ungroup() %>% dplyr::select(STRATA_FULL,
                                                                                                                                          DISCARD_SOURCE 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>%# top_n(10) %>%
		DT::datatable(caption = 'Discard rates for state trips by strata for Red Hake')
```

```{r silver hake federal discard table }

db_silver %>% 	
  filter(DISCARD_SOURCE != 'O' & FED_OR_STATE=="FED") %>%
  group_by(STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
   ungroup() %>% dplyr::select(STRATA_FULL,
                                                                                                                                          DISCARD_SOURCE 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>%# top_n(10) %>%
		DT::datatable(caption = 'Discard rates for federal trips by strata for Silver Hake')
```

```{r silver hake state discard table }

db_silver %>% 	
  filter(DISCARD_SOURCE != 'O' & FED_OR_STATE=="STATE") %>%
  group_by(STRATA_FULL, DISCARD_SOURCE, FED_OR_STATE) %>%
  dplyr::summarise(SUBTRIP_KALL = sum(SUBTRIP_KALL)
                       , CAMS_DISCARD_RATE = round(mean(CAMS_DISCARD_RATE), digits=5)
                       , DISC_MORT_RATIO = mean(DISC_MORT_RATIO)
                      # , CV = mean(CV, na.rm=TRUE)
                       , DISCARD = round(sum(CAMS_DISCARD))) %>%
                      # , N_OBS_TRIPS_F = mean(N_OBS_TRIPS_F)) %>% 
   ungroup() %>% dplyr::select(STRATA_FULL,
                                                                                                                                          DISCARD_SOURCE 
             # ,N_OBS_TRIPS_F
              ,CAMS_DISCARD_RATE
              ,SUBTRIP_KALL
              ,DISC_MORT_RATIO
              ,DISCARD) %>%
             # ,CV) 
             dplyr::arrange(desc(DISCARD)) %>%# top_n(10) %>%
		DT::datatable(caption = 'Discard rates for state trips by Strata for Silver Hake')
```

```{r dogfish gear fed discard table }
db_dogfish %>% filter(FED_OR_STATE == 'FED') %>% group_by(DISCARD_SOURCE, CAMS_GEAR_GROUP, SPECIES_STOCK) %>% dplyr::summarise(D = sum(CAMS_DISCARD), K = sum(SUBTRIP_KALL), Drate = mean(CAMS_DISCARD_RATE), Dmort = mean(DISC_MORT_RATIO)) %>% arrange(desc(D))%>% top_n(10) %>% dplyr::mutate(across(where(is.numeric), round, 4)) %>% mutate(across(ends_with("D"), round, 0))%>%
		DT::datatable(caption = 'Discard totals for federal trips by stock, gear and discard source for Dogfish')
```

```{r dogfish gear state discard table }
db_dogfish %>% filter(FED_OR_STATE == 'STATE') %>% group_by(DISCARD_SOURCE, CAMS_GEAR_GROUP, SPECIES_STOCK) %>% dplyr::summarise(D = sum(CAMS_DISCARD), K = sum(SUBTRIP_KALL), Drate = mean(CAMS_DISCARD_RATE), Dmort = mean(DISC_MORT_RATIO)) %>% arrange(desc(D))%>% top_n(10) %>% dplyr::mutate(across(where(is.numeric), round, 4))%>% mutate(across(ends_with("D"), round, 0))%>%
		DT::datatable(caption = 'Discard totals for state trips by stock, gear and discard source for Dogfish')
```

```{r monkfish gear fed discard table }
db_monkfish %>% filter(FED_OR_STATE == 'FED') %>% group_by(DISCARD_SOURCE, CAMS_GEAR_GROUP, SPECIES_STOCK) %>% dplyr::summarise(D = sum(CAMS_DISCARD), K = sum(SUBTRIP_KALL), Drate = mean(CAMS_DISCARD_RATE), Dmort = mean(DISC_MORT_RATIO)) %>% arrange(desc(D))%>% top_n(10) %>% dplyr::mutate(across(where(is.numeric), round, 4))%>% mutate(across(ends_with("D"), round, 0))%>%
		DT::datatable(caption = 'Discard totals for federal trips by stock, gear and discard source for Monkfish')
```

```{r monkfish gear state discard table }
db_monkfish %>% filter(FED_OR_STATE == 'STATE') %>% group_by(DISCARD_SOURCE, CAMS_GEAR_GROUP, SPECIES_STOCK) %>% dplyr::summarise(D = sum(CAMS_DISCARD), K = sum(SUBTRIP_KALL), Drate = mean(CAMS_DISCARD_RATE), Dmort = mean(DISC_MORT_RATIO)) %>% arrange(desc(D))%>% top_n(10) %>% dplyr::mutate(across(where(is.numeric), round, 4))%>% mutate(across(ends_with("D"), round, 0))%>%
		DT::datatable(caption = 'Discard totals for state trips by stock, gear and discard source for Monkfish')
```

```{r offshorehake gear fed discard table }
db_offshore %>% filter(FED_OR_STATE == 'FED') %>% group_by(DISCARD_SOURCE, CAMS_GEAR_GROUP, SPECIES_STOCK) %>% dplyr::summarise(D = sum(CAMS_DISCARD), K = sum(SUBTRIP_KALL), Drate = mean(CAMS_DISCARD_RATE), Dmort = mean(DISC_MORT_RATIO)) %>% arrange(desc(D))%>% top_n(10) %>% dplyr::mutate(across(where(is.numeric), round, 4))%>% mutate(across(ends_with("D"), round, 0))%>%
		DT::datatable(caption = 'Discard totals for federal trips by stock, gear and discard source for Offshorehake')
```

```{r offshorehake gear state discard table }
db_offshore %>% filter(FED_OR_STATE == 'STATE') %>% group_by(DISCARD_SOURCE, CAMS_GEAR_GROUP, SPECIES_STOCK) %>% dplyr::summarise(D = sum(CAMS_DISCARD), K = sum(SUBTRIP_KALL), Drate = mean(CAMS_DISCARD_RATE), Dmort = mean(DISC_MORT_RATIO)) %>% arrange(desc(D))%>% top_n(10) %>% dplyr::mutate(across(where(is.numeric), round, 4))%>% mutate(across(ends_with("D"), round, 0))%>%
		DT::datatable(caption = 'Discard totals for state trips by stock, gear and discard source for Offshorehake')
```

```{r redhake gear fed discard table }
db_red %>% filter(FED_OR_STATE == 'FED') %>% group_by(DISCARD_SOURCE, CAMS_GEAR_GROUP, SPECIES_STOCK) %>% dplyr::summarise(D = sum(CAMS_DISCARD), K = sum(SUBTRIP_KALL), Drate = mean(CAMS_DISCARD_RATE), Dmort = mean(DISC_MORT_RATIO)) %>% arrange(desc(D))%>% top_n(10) %>% dplyr::mutate(across(where(is.numeric), round, 4))%>% mutate(across(ends_with("D"), round, 0))%>%
		DT::datatable(caption = 'Discard totals for federal trips by stock, gear and discard source for Redhake')
```

```{r redhake gear state discard table }
db_red %>% filter(FED_OR_STATE == 'STATE') %>% group_by(DISCARD_SOURCE, CAMS_GEAR_GROUP, SPECIES_STOCK) %>% dplyr::summarise(D = sum(CAMS_DISCARD), K = sum(SUBTRIP_KALL), Drate = mean(CAMS_DISCARD_RATE), Dmort = mean(DISC_MORT_RATIO)) %>% arrange(desc(D))%>% top_n(10) %>% dplyr::mutate(across(where(is.numeric), round, 4))%>% mutate(across(ends_with("D"), round, 0))%>%
		DT::datatable(caption = 'Discard totals for state trips by stock, gear and discard source for Redhake')
```

```{r silverhake gear fed discard table }
db_silver %>% filter(FED_OR_STATE == 'FED') %>% group_by(DISCARD_SOURCE, CAMS_GEAR_GROUP, SPECIES_STOCK) %>% dplyr::summarise(D = sum(CAMS_DISCARD), K = sum(SUBTRIP_KALL), Drate = mean(CAMS_DISCARD_RATE), Dmort = mean(DISC_MORT_RATIO)) %>% arrange(desc(D))%>% top_n(10) %>% dplyr::mutate(across(where(is.numeric), round, 4))%>% mutate(across(ends_with("D"), round, 0))%>%
		DT::datatable(caption = 'Discard totals for federal trips by stock, gear and discard source for Silverhake')
```

```{r silverhake gear state discard table }
db_silver %>% filter(FED_OR_STATE == 'STATE') %>% group_by(DISCARD_SOURCE, CAMS_GEAR_GROUP, SPECIES_STOCK) %>% dplyr::summarise(D = sum(CAMS_DISCARD), K = sum(SUBTRIP_KALL), Drate = mean(CAMS_DISCARD_RATE), Dmort = mean(DISC_MORT_RATIO)) %>% arrange(desc(D))%>% top_n(10) %>% dplyr::mutate(across(where(is.numeric), round, 4))%>% mutate(across(ends_with("D"), round, 0))%>%
		DT::datatable(caption = 'Discard totals for state trips by stock, gear and discard source for Silverhake')
```
### Appendix

For more information regarding stratification variables and discard rates by species see

MAPS.CAMS_STATAREA_STOCK
MAPS.CAMS_DISCARD_MORTALITY_STOCK
MAPS.CAMS_GEARCODE_STRATA
MAPS.CAMS_MASTER_GEAR
MAPS.CAMS_OBS_CATCH
